```{r set-knitr-options, cache=FALSE, echo=FALSE}
library("knitr")
opts_chunk$set(cache=TRUE, echo=FALSE, message=FALSE, autodep=TRUE,
  results='hide')
opts_chunk$set(warning=FALSE, message=FALSE, cache.path="../cache/",
  cache.lazy=TRUE)
```
```{r get-dat, cache=FALSE}
load("../data-generated/output-values.rda")
h1 <- readRDS("../data-generated/h1_effect.rds")
spoke <- readRDS("../data-generated/spoke_fraction.rds")
g2_dens <- readRDS("../data-generated/g2_prob_dens_above_zero.rds")
```

# Abstract

<!-- Fishing is inherently risky and we have long-sought regulations and fishing behaviors that reduce this risk. -->
<!-- The increasing the diversity of species fished has been proposed to reduce risk for fishermen -->

Individuals relying on natural resources as their primary source of income
often experience high levels of income variability
due to the combined effects of environmental, biological, and price variation.
Commercial fishers are a prime example and
recent work has sought to identify regulations and fishing behaviours
that can reduce this variability.
One candidate has been that fishing higher levels of species diversity
decreases income variability.
However, it is unknown if this species-diversity effect
on income variability holds
across multiple approaches to diversifying catch,
at the level of individual permit holders,
and across different fishing strategies.
Here, we simultaneously estimate the effects of three
methods of diversification for Alaskan fishery permit holders
on revenue and revenue variability:
(1) owning multiple permits,
(2) owning permits that inherently fish higher species diversity,
and (3) increasing species diversity within the constraints of a permit set.
We control for changes in effort and other
latent effects through time.
We find that
fishing an additional permit is associated with higher revenue
and lower variability (e.g.,
`r 100-spoke[[2]]`% [`r 100-spoke[[3]]`--`r 100-spoke[[1]]`% IQR] lower
expected standard deviation
for halibut plus another permit vs. single halibut permits);
fishing permits with higher species diversity reduces variability
(`r h1[[2]]`% [`r h1[[1]]`--`r h1[[3]]`% 95% CI] reduction per unit
increase in mean effective species diversity);
but that increasing species diversity within
the constraints of a permit set has a fishery-dependent effect on revenue and
is usually (`r g2_dens`% probability)
associated with increased uncertainty in revenue.
At the same time, Alaskan commercial fisheries
have been increasingly focussed on
single-permit (often salmon) fishing strategies
with single-permit fishers earning
more than multiple-permit fishers
for the first time in 2013.
Our results suggest that the benefit of species diversity
for revenue variability
is mostly an across-permit effect, not
a benefit that can be easily harnessed by an individual permit holder.
Effective methods of diversification ---
purchasing additional permits or switching permits ---
are costly and often limited by regulations such as limited entry
programs and individual transferable quotas (ITQs).
Take home ...

\bigskip
\noindent
Extra:

- there is often limited room for intentional diversification within a permit
- in fact, specializing can be profitable within some fisheries

\noindent
Main message:
Diversified portfolios of fishing permits and species reduce risk for fishermen,
but changing species diversity can be risky, costly, and limited by regulations.

<!-- , which is concerning  -->
<!-- given that some fishing regulations and market forces -->
<!-- may be increasingly promoting specialization. -->
<!-- Main message in 25 words or less: -->

<!-- Take 1: -->
<!-- - There is a trend towards specializing on single salmon permits in Alaska -->
<!-- - Diversified portfolios of fishing permits and species reduce risk, -->
<!--   but changing species diversity for an individual permit holder can be risky -->
<!-- - The types of diversity needed to reduce risk are often restricted by -->
<!--   cost and regulations -->

<!-- Take 2: -->
<!-- - The effect of species diversification on revenue and risk for fishermen -->
<!-- depends on the type of diversification and the fishery. -->

<!-- Introduction, Methods, Results, Discussion, Conclusion:
<!-- 600, 900, 500, 800, 200, -->

<!-- write out each paragraph’s main message in 15 words or less -->
<!-- circulate skeleton ... this is the time to fix problems with flow, -->
<!-- ideas/content, main message -->

# Introduction

Natural resources such as forests, land animals, and fishes
provide a primary source of income for millions of people worldwide [REFS].
However, environmental,
biological, and price variation
combined with regulatory changes
and human-caused environmental disasters
generate high levels of income variability and risk
for these individuals.
For example, ...
And EVOS example

\bigskip

Targeting a higher diversity of species
has been proposed as a risk-buffering mechanism...
Give examples from other sectors

\bigskip

This effect has been identified for fisheries.
At community level, ... [@sethi2014]
At vessel level, ... [@kasperski2013] ...

\bigskip

Despite this existing work,
key questions remain on the relationship between species diversity and
fishing revenue variability.
First, the effect of changing species diversity
for individual permit holders on a year-by-year basis is unknown ---
existing approaches have worked at the vessel or community level
and condensed time series into single metrics of variability or risk.
Second, the effect of changing species diversity
on revenue itself is unknown ---
existing approaches have focused solely on variability or risk.
Third, it is unknown if the approach to
diversifying species fished affects the relationship between
species diversity and revenue variability ---
a fisheries permit holder can purchase
an additional permit to target other species,
pick a permit that inherently targets many species,
or increase species diversity
within the constraints of his or her existing permits.
Many permit holders may be limited to the last approach
given that the first two may be costly and limited by regulations,
but it is unclear if the benefits and costs are the same.
Here, we address these questions by
simultaneously estimating the effects of
these three approaches to species diversification
for Alaskan fishery permit holders on
annual revenue and revenue variability.
Our model makes use of the entire time series,
controls for effort and latent time effects,
and allows the effect of species diversity
to vary by fishing strategy (sets of similar permits)
in a hierarchical model.




<!-- Multiple ways one can diversify: -->
<!-- buy another permit, pick diverse permit, diversify within existing constraints. -->


# Methods

- overview

## Data

We obtained landings and revenue data for
all permit holders in Alaska
from `r range_yrs[1]` to `r range_yrs[2]`
from the Alaskan Commercial Fisheries Entry Commission (CFEC).
These data represent a total of
nearly `r round(ndat$initial_tot_rev/1e9)` billion 2009 USD
and a combined `r round(ndat$initial/1e6,1)` million rows of data.
We adjusted for inflation by setting all revenues in 2009 USD
based on data from TODO.
We implemented a number of filtering steps
to focus on individuals actively engaged in commercial fishing
(Supporting Information).
In particular, we removed permit holders whose median revenue
was less than $10,000 2009 USD.

Our models focused on the predictor species diversity,
while controlling for effort.
We calculated effective species diversity for
permit holder $i$ and year $t$ as the inverse of
Simpson's diversity index $\lambda$ [@simpson1949]
weighted by revenue $R$ ('species diversity'):

\begin{equation}
1/\lambda_{i,t} = 1 / \left(\sum\limits_{s=1}^{n_{i,t}}
  \left(R_{i,s,t}/ \sum\limits_{s=1}^{n_{i,t}} R_{i,s,t}\right)\right)^2,
\end{equation}

\noindent
where $s$ indexes species from $1$ through $n$.
We calculated effort or effective season length as
the sum of days fished across each permit each year.
For example, if someone fished for 30 days
on a salmon drift gillnet permit in Southeast Alaska
and 40 days on a statewide longline halibut permit,
that individual's season length
would be 70 days.

## Hierarchical model

The basic model is a hierarchical linear regression with Gaussian errors
predicting $\log$ revenue, $\log(R_{i,t})$, for
each fisher $i$ and time (year) $t$:

\begin{equation}
\log(R_{i,t}) \sim \mathrm{N} \left(\widehat{\log(R_{i,t})}, \widehat{\sigma_{i,t}^2} \right).
\end{equation}

\noindent
Each year $t$ of revenue for fisher $i$
is assigned a strategy indexed by $j$ and
expected $\log(R_{i,t})$ is modelled as:

\begin{equation}
\label{eq:mean}
\widehat{\log(R_{i,t})} = \beta_{0,j,t} -
  \beta_{1,j} S_{i,t} I_{i,t} +
  \beta_{2,j} S_{i,t} I_{i,t} +
  \beta_3 D_{i,t} -
  \beta_4 S_{i,t} D_{i,t} I_{i,t} +
  \beta_5 S_{i,t} D_{i,t} I_{i,t} +
  \log(R_{i,t-1})
\end{equation}

\noindent
where $S_{i,t}$ is the percent change in species diversity,
$\log \left(\mathrm{div}_{i,t}/\mathrm{div}_{i,t-1}\right)$,
$D_{i,t}$ is the percent change in days fished,
$\log \left(\mathrm{days}_{i,t}/\mathrm{days}_{i,t-1}\right)$, and
$I_{i,t}$ is an indicator variable that takes a value $0$
if $\mathrm{div}_{i,t} \ge \mathrm{div}_{i,t-1}$
(species diversity increasing; generalizing)
and takes a value $1$ if
$\mathrm{div}_{i,t} < \mathrm{div}_{i,t-1}$
(species diversity decreasing; specializing).

The term $\log(R_{i,t-1}$) is treated as an offset term:
it is added on the right side of the equation
to essentially model percent change in revenue,
i.e., $\log(R_{i,t}) - \log(R_{i,t-1}) = \log(R_{i,t}/R_{i,t-1})$.
While this approach required filtering the data
to remove person-year combinations without a subsequent year of revenue,
it has a number of advantages:
it eliminates having to deal with the non-stationary
time series and eliminates having to
estimate thousands of people-specific intercepts.

The $\beta$'s represent estimated coefficients, some of which are
allowed to vary.
The intercepts $\beta_{0,j,t}$ are allowed to vary by
strategy-year combinations,
are centered on zero, and are constrained by a normal distribution:

\begin{equation}
\beta_{0,j,t} \sim \mathrm{N} \left(0, \sigma^2_{\beta_{j,t}} \right).
\end{equation}

\noindent
This, combined with the lack of a global intercept, constrains
the mean change in log revenue across all fishers and years to
be zero. I.e., we assume that the data are stationary
after accounting for all covariates and including
an intercept that varies with strategy-year combinations.
Initial versions of the model
confirmed that the global intercept along with stragey-level
intercepts would be estimated at almost exactly zero.

The slopes $\beta_{1,j}$ and $\beta_{1,j}$
are then allowed to vary by strategy with estimated means:

\begin{align}
\beta_{1,j} &\sim \mathrm{N} \left(\beta_1, \sigma^2_{\beta_1}\right),\\
\beta_{2,j} &\sim \mathrm{N} \left(\beta_1, \sigma^2_{\beta_2}\right).
\end{align}

\noindent
The $\beta_{1,j}$'s represent how much an X\% decrease in species
diversity within a strategy translates to a Y\% change in revenue.
The $\beta_{2,j}$'s represent how much an X\% increase in species
diversity within a strategy translates to a Y\% change in revenue.
These are after controlling for percent change in days fished, $\beta_3$,
and the interaction between days fished and percent change in species
diversity, $\beta_4$ and $\beta_5$.

Whereas a typical hierarchical linear regression model
would estimate a single residual error scale term, $\sigma$,
we model the scale of the residual error  with
another hierarchical model with a similar form:

\begin{equation}
\label{eq:cv}
\widehat{\sigma_{i,t}} =
\exp \left( \gamma_{0,j} -
\gamma_{1,j} S_{i,t} I_{i,t} +
\gamma_{2,j} S_{i,t} I_{i,t} +
\gamma_3 D_{i,t} -
\gamma_4 S_{i,t} D_{i,t} I_{i,t} +
\gamma_5 S_{i,t} D_{i,t} I_{i,t}\right).
\end{equation}

\noindent
Again, $S_{i,t}$ and $I_{i,t}$ represent, respectively,
percent change in species
diversity and an indicator variable for increasing or
decreasing species diversity.
We exponentiate the equation to ensure all scale parameters, $\sigma_{i,t}$,
are positive. The $\gamma$'s are estimated coefficients, three of which
are allowed to vary by strategy, constrained by normal distributions:

\begin{align}
\gamma_{0,j}  &\sim \mathrm{N} \left(\gamma_0 + \eta_1 M_j + \eta_2 E_j, \sigma^2_{\gamma_0}\right),\\
\gamma_{1,j}  &\sim \mathrm{N} \left(\gamma_1, \sigma^2_{\gamma_1}\right),\\
\gamma_{2,j}  &\sim \mathrm{N} \left(\gamma_2, \sigma^2_{\gamma_2}\right).
\end{align}

\noindent
The mean of the $\gamma_{0,j}$ coefficients is
modelled with
two strategy-level predictors.
The symbols $M_j$ and $E_j$ represent the raw mean species diversity and
mean combined fishing season length (effort) for strategy $j$.
The coefficients $\eta_1$ and $\eta_2$ represent predictors
in this strategy-level regression
estimating effects across strategies.

We fit our models with Stan 2.11.1
[@hoffman2014; @standevelopmentteam2015; @standevelopmentteam2015a; @carpenter2015]
and R 3.3.1 [@rcoreteam2016].
Stan implements the
No-U-Turn Hamiltonian Markov chain Monte Carlo algorithm [@hoffman2014]
to perform Bayesian statistical inference.
We ran four chains and 2,000 iterations and discarded
the first 1,000 iterations of each chain
as warm up (i.e. 4,000 final samples).
We checked for chain convergence visually with trace plots.
We also ensured that $\hat{R}<1.05$ (the potential scale reduction factor)
and that the effective sample size was
greater than 200 for all parameters [@gelman2014].
We verified that our model was coded correctly by
comparing our estimates to an ad-hoc two-stage
model fit with the lme4 R package [@bates2015],
and by fitting our model to simulated data
and recovering the true values.
Code to reproduce our results (with a simulated data set)
is available at <https://github.com/NCEAS/pfx-commercial>.

# Results

# Discussion

\bibliographystyle{apalike}
\bibliography{refs}

\clearpage

# Figures


\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.7\textwidth]{../figs/fig1-jk.pdf}
\caption{
Trends in revenue and diversification over time.
(A) Map of fishing regions in Alaska, USA.
(B) Total revenue in billions of 2009 U.S. dollars from `r range_yrs[1]` to
`r range_yrs[2]` for multiple (grey) and single (blue) permit holders.
(C) Proportion of single permit holders for all fisheries (thick black line)
and with permit holders omitted each year who fished any of the
`r n_jacknifed` top species groups (thin lines).
(D) Mean effective diversity of species fished (Simpson's diversity index)
for all fisheries (thick black line)
and with permit holders omitted each year who fished any of the
`r n_jacknifed` top species groups (thin lines).
}
\label{fig:ts}
\end{center}
\end{figure}

\clearpage

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.9\textwidth]{../figs/stan-gg-spoke2.pdf}
\caption{Diversifying permits reduces risk and usually increases revenue.
Two examples of expected revenue variability
and median revenue for
owners of (A) individual halibut and (B) gillnet herring roe permits
(dark blue)
compared with owners of the same permit plus one more (medium blue)
or two more permits (light blue).
On the y-axis, dots and lines represent posterior median and 95\%
credible intervals of expected standard deviation.}
\label{fig:spokes}
\end{center}
\end{figure}

\clearpage

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.7\textwidth]{../figs/stan-across-strategy-variability2.pdf}
\caption{Individuals who fish
permits and permit combinations (permit strategies) with higher species diversity
have lower expected revenue variability.
On the y-axis, dots, thick lines, and thin lines represent posterior medians,
50\%, and 95\% credible intervals for
the `r nstr` most common permit strategies.
Thick black line, and dark and light grey shaded regions
indicate median, 50\%, and 95\% credible intervals
of strategy-level regression built into the hierarchical model.
Estimated variability represents expected standard deviation
for a permit holder who does not change
species diversity or days fished from year to year
with the strategy-level effect of mean days fished removed.
}
\label{fig:across}
\end{center}
\end{figure}

\clearpage

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=1.1\textwidth]{../figs/stan-offset-break-anti-spaghetti.pdf}
\caption{Effects of (A) generalizing and (B) specializing from year to year for
individual permit holders across the `r nstr` most common strategies.
Changing species diversity from year to year
is associated with higher variability
for individuals within the strategies above the zero line.
Generalizing (A) or specializing (B) is associated with
greater expected revenue to the right of the zero line
and lower expected revenue to the left of the zero line.
These effects are estimated holding
fishing season length constant from year to year.
Dots and line segments represent medians and
50\% credible intervals of posteriors.
}
\label{fig:within}
\end{center}
\end{figure}

\clearpage

# Notes

Fishing can be big $ but is risky, for example, Alaska


What's been done

## @kasperski2013
- vessel-level CFEC data
- dome-shaped result
- Herfindahl-Hirschman Index/HHI (Simpson)
- diversity decreasing for those still fishing in 2010, but not otherwise

## @sethi2012b
- risk measures to 90 Alaskan fishing fishery/fleets
- revenue more variable than catch (+ price = extra variability)
- large catches, big fleets = lower risk

## @sethi2010b
- risk management for fisheries review

## @sethi2012
- equal-share catch-pooling cooperatives in king and snow crab
- "variability reduction depends on the degree of correlation amongst
  participants’ catches"

## @sethi2014
- community-level CFEC data
- number of fisheries strongest predictor of (lower) CV/CVaR
- also simpson index
- suggest barriers to commercial entry increase community risk by restricting
  access to 'variety of fisheries'
- some communities have limited ability to diversify; may need to diversify
  outside of fishing

[@perruso2005]
[@hilborn2001]
[@edwards2004]
[@trenkel2013]
[@mumford2009]
[@oostenbrugge2002]
[@baldursson1997]
[@jardine2015]
[@sanchirico2008]
