```{r set-knitr-options, cache=FALSE, echo=FALSE}
library("knitr")
opts_chunk$set(cache=TRUE, echo=FALSE, message=FALSE, autodep=TRUE,
  results='hide')
opts_chunk$set(warning=FALSE, message=FALSE, cache.path="../cache/",
  cache.lazy=TRUE)
```
```{r get-dat, cache=FALSE}
load("../data-generated/output-values.rda")
h1 <- readRDS("../data-generated/h1_effect.rds")
spoke <- readRDS("../data-generated/spoke_fraction.rds")
g2_dens <- readRDS("../data-generated/g2_prob_dens_above_zero.rds")
n_grouped_permits <- readRDS("../data-generated/n_grouped_permits.rds")
```

# Abstract

<!-- Fishing is inherently risky and we have long-sought regulations and fishing behaviors that reduce this risk. -->
<!-- The increasing the diversity of species fished has been proposed to reduce risk for fishermen -->

Individuals relying on natural resources as their primary source of income
often experience high levels of income variability
due to the combined effects of environmental, biological, and price variation.
Commercial fishers are a prime example and
recent work has sought to identify regulations and fishing behaviours
that can reduce this variability.
One candidate has been that fishing higher levels of species diversity
decreases income variability.
However, it is unknown if this species-diversity effect
on income variability holds
across multiple approaches to diversifying catch,
at the level of individual permit holders,
and across different fishing strategies.
Here, we simultaneously estimate the effects of three
methods of diversification for Alaskan fishery permit holders
on revenue and revenue variability:
(1) owning multiple permits,
(2) owning permits that inherently fish higher species diversity,
and (3) increasing species diversity within the constraints of a permit set.
We control for changes in effort and other
latent effects through time.
We find that
fishing an additional permit is associated with higher revenue
and lower variability (e.g.,
`r 100-spoke[[2]]`% [`r 100-spoke[[3]]`--`r 100-spoke[[1]]`% IQR] lower
expected standard deviation
for halibut plus another permit vs. single halibut permits);
fishing permits with higher species diversity reduces variability
(`r h1[[2]]`% [`r h1[[1]]`--`r h1[[3]]`% 95% CI] reduction per unit
increase in mean effective species diversity);
but that increasing species diversity within
the constraints of a permit set has a fishery-dependent effect on revenue and
is usually (`r g2_dens`% probability)
associated with increased uncertainty in revenue.
At the same time, Alaskan commercial fisheries
have been increasingly focussed on
single-permit (often salmon) fishing strategies
with single-permit fishers earning
more than multiple-permit fishers
for the first time in 2013.
Our results suggest that the benefit of species diversity
for revenue variability
is mostly an across-permit effect, not
a benefit that can be easily harnessed by an individual permit holder.
Effective methods of diversification ---
purchasing additional permits or switching permits ---
are costly and often limited by regulations such as limited entry
programs and individual transferable quotas (ITQs).
Take home ...

\bigskip
\noindent
Extra:

- there is often limited room for intentional diversification within a permit
- in fact, specializing can be profitable within some fisheries

\noindent
Main message:
Diversified portfolios of fishing permits and species reduce risk for fishermen,
but changing species diversity can be risky, costly, and limited by regulations.

<!-- , which is concerning  -->
<!-- given that some fishing regulations and market forces -->
<!-- may be increasingly promoting specialization. -->
<!-- Main message in 25 words or less: -->

<!-- Take 1: -->
<!-- - There is a trend towards specializing on single salmon permits in Alaska -->
<!-- - Diversified portfolios of fishing permits and species reduce risk, -->
<!--   but changing species diversity for an individual permit holder can be risky -->
<!-- - The types of diversity needed to reduce risk are often restricted by -->
<!--   cost and regulations -->

<!-- Take 2: -->
<!-- - The effect of species diversification on revenue and risk for fishermen -->
<!-- depends on the type of diversification and the fishery. -->

<!-- Introduction, Methods, Results, Discussion, Conclusion:
<!-- 600, 900, 500, 800, 200, -->

<!-- write out each paragraph’s main message in 15 words or less -->
<!-- circulate skeleton ... this is the time to fix problems with flow, -->
<!-- ideas/content, main message -->

# Introduction

Natural resources such as forests, land animals, and fishes
provide a primary source of income for millions of people worldwide [REFS].
However, environmental,
biological, and price variation
combined with regulatory changes
and human-caused environmental disasters
generate high levels of income variability and risk
for these individuals.
For example, ...
And EVOS example

\bigskip

Targeting a higher diversity of species
has been proposed as a risk-buffering mechanism...
Give examples from other sectors

\bigskip

This effect has been identified for fisheries.
At community level, ... [@sethi2014]
At vessel level, ... [@kasperski2013] ...

\bigskip

Despite this existing work,
key questions remain on the relationship between species diversity and
fishing revenue variability.
First, the effect of changing species diversity
for individual permit holders on a year-by-year basis is unknown ---
existing approaches have worked at the vessel [@kasperski2013],
fleet [@sethi2012b], or community level [@sethi2014],
and condensed time series into single metrics of variability or risk.
Second, the effect of changing species diversity
on revenue itself is unknown ---
existing approaches have focused solely on variability or risk.
Third, it is unknown if the approach to
diversifying species fished affects the relationship between
species diversity and revenue variability ---
a fishery permit holder can purchase
an additional permit to target other species,
pick a permit that inherently targets many species,
or increase species diversity
within the constraints of his or her existing permits.
Many permit holders may be limited to the last approach
given that the first two may be costly and limited by regulations,
but it is unclear if the benefits and costs are the same.
Here, we address these questions by
simultaneously estimating the effects of
these three species diversification approaches
on annual revenue and revenue variability
for Alaskan fishery permit holders.
Our model makes use of entire time series,
controls for effort and latent time effects,
and allows the effect of species diversity
to vary by fishing strategy (sets of similar permits)
in a hierarchical model.




<!-- Multiple ways one can diversify: -->
<!-- buy another permit, pick diverse permit, diversify within existing constraints. -->


# Methods

Overview paragraph here.

## Data

We obtained fisheries landings and revenue data for
all permit holders in Alaska
from `r range_yrs[1]` to `r range_yrs[2]`
from the Alaskan Commercial Fisheries Entry Commission (CFEC).
These data represent a total of
nearly `r round(ndat$initial_tot_rev/1e9)` billion 2009 USD
and a combined `r round(ndat$initial/1e6,1)` million rows of data.
We adjusted for inflation by setting all revenues in 2009 USD
based on data from TODO.
We implemented a number of filtering steps
to focus on individuals actively engaged in commercial fishing
(Supporting Information).
In particular, we removed permit holders whose median revenue
was less than $10,000 2009 USD.

We combined similar permits into permit groupings 
and identified sets of 
permit groupings held by an individual 
in a specific year as 'strategies' (Supporting Information). 
For example, we combined permits that were the same except for vessel size,
and we combined a number of permits that were targeting a single species 
and differed only in region fished.
Therefore, someone fishing herring roe in Kodiak
and someone fishing herring roe in Alaska Peninsula
were both considered to have a 'herring roe' permit group.
For salmon permits, 
we grouped permits using the same gear type across regions
for groups that shared similar species composition. 
After this process of combining permits,
we were left with `r n_grouped_permits`
unique fishing permit groups (Table S\ref{tab:strategies}),
which we aggregated
within each person-year combination 
to form permit strategies.
For example, if someone fished halibut
and sablefish in the same year,
their strategy for that year would be `halibut-sablefish'.
Specific details on our strategy definitions are available in 
the Supporting Information.

Our models focused on species diversity
as a possible predictor of revenue and revenue variability
while controlling for effort.
We calculated effective species diversity for
permit holder $i$ and time (year) $t$ as the inverse of
Simpson's diversity index $\lambda$ [@simpson1949]
weighted by revenue $R$ ('species diversity'):

\begin{equation}
1/\lambda_{i,t} = 1 / \left(\sum\limits_{s=1}^{n_{i,t}}
  \left(R_{i,s,t}/ \sum\limits_{s=1}^{n_{i,t}} R_{i,s,t}\right)\right)^2,
\end{equation}

\noindent
where $s$ indexes species from $1$ through $n$.
We calculated effort or effective season length as
the sum the season lengths for each permit each year.
Season length for a given permit 
was calculated as the the number of days between 
the first instance of fishing a permit 
and the last instance of fishing a permit each year.
For example, if someone fished for 30 days
on a salmon drift gillnet permit in Southeast Alaska
and 40 days on a statewide longline halibut permit,
that individual's season length
would be 70 days.

## Hierarchical model

The basic form of our model was a 
hierarchical linear regression with Gaussian errors
predicting $\log$ revenue, $\log(R_{i,t})$, for
each fisher $i$ and time $t$:

\begin{equation}
\log(R_{i,t}) \sim \mathrm{N} \left(\widehat{\log(R_{i,t})}, \widehat{\sigma_{i,t}^2} \right).
\end{equation}

\noindent
Each year $t$ of revenue for fisher $i$
is assigned a strategy indexed by $j$ 
(Supporting Information, Table \ref{tab:strategies})
and expected $\log(R_{i,t})$ is modelled as:

\begin{equation}
\label{eq:mean}
\widehat{\log(R_{i,t})} = \beta_{0,j,t} -
  \beta_{1,j} S_{i,t} I_{i,t} +
  \beta_{2,j} S_{i,t} I_{i,t} +
  \beta_3 D_{i,t} -
  \beta_4 S_{i,t} D_{i,t} I_{i,t} +
  \beta_5 S_{i,t} D_{i,t} I_{i,t} +
  \log(R_{i,t-1}),
\end{equation}

\noindent
where $S_{i,t}$ is the percent change in species diversity,
$\log \left(\mathrm{div}_{i,t}/\mathrm{div}_{i,t-1}\right)$,
$D_{i,t}$ is the percent change in days fished,
$\log \left(\mathrm{days}_{i,t}/\mathrm{days}_{i,t-1}\right)$, and
$I_{i,t}$ is an indicator variable that takes a value $0$
if $\mathrm{div}_{i,t} \ge \mathrm{div}_{i,t-1}$
(species diversity increasing; generalizing)
and takes a value $1$ if
$\mathrm{div}_{i,t} < \mathrm{div}_{i,t-1}$
(species diversity decreasing; specializing).

The term $\log(R_{i,t-1}$) is treated as an offset term:
it is added on the right side of the equation
to essentially model percent change in revenue,
i.e., $\log(R_{i,t}) - \log(R_{i,t-1}) = \log(R_{i,t}/R_{i,t-1})$.
While this approach required filtering the data
to remove person-year combinations without a subsequent year of revenue,
it has a number of advantages:
it eliminates having to deal with the non-stationary
time series and eliminates having to
estimate thousands of people-specific intercepts.

The $\beta$'s represent estimated coefficients, some of which are
allowed to vary.
The intercepts $\beta_{0,j,t}$ are allowed to vary by
strategy-year combinations,
are centered on zero, and are constrained by a normal distribution:

\begin{equation}
\beta_{0,j,t} \sim \mathrm{N} \left(0, \sigma^2_{\beta_{j,t}} \right).
\end{equation}

\noindent
This, combined with the lack of a global intercept, constrains
the mean change in log revenue across all fishers and years to
be zero. I.e., we assume that the data are stationary
after accounting for all covariates and including
an intercept that varies with strategy-year combinations.
Initial versions of the model
confirmed that the global intercept along with stragey-level
intercepts would be estimated at almost exactly zero.

The slopes $\beta_{1,j}$ and $\beta_{1,j}$
are then allowed to vary by strategy with estimated means:

\begin{align}
\beta_{1,j} &\sim \mathrm{N} \left(\mu_{\beta_1}, \sigma^2_{\beta_1}\right),\\
\beta_{2,j} &\sim \mathrm{N} \left(\mu_{\beta_2}, \sigma^2_{\beta_2}\right).
\end{align}

\noindent
The $\beta_{1,j}$'s represent how much an X\% decrease in species
diversity within a strategy translates to a Y\% change in revenue.
The $\beta_{2,j}$'s represent how much an X\% increase in species
diversity within a strategy translates to a Y\% change in revenue.
These are after controlling for percent change in days fished, $\beta_3$,
and the interaction between days fished and percent change in species
diversity, $\beta_4$ and $\beta_5$.

Whereas a typical hierarchical linear regression model
would estimate a single residual error scale term, $\sigma$,
we model the scale of the residual error  with
another hierarchical model with a similar form:

\begin{equation}
\label{eq:cv}
\widehat{\sigma_{i,t}} =
\exp \left( \gamma_{0,j} -
\gamma_{1,j} S_{i,t} I_{i,t} +
\gamma_{2,j} S_{i,t} I_{i,t} +
\gamma_3 D_{i,t} -
\gamma_4 S_{i,t} D_{i,t} I_{i,t} +
\gamma_5 S_{i,t} D_{i,t} I_{i,t}\right).
\end{equation}

\noindent
Again, $S_{i,t}$ and $I_{i,t}$ represent, respectively,
percent change in species
diversity and an indicator variable for increasing or
decreasing species diversity.
We exponentiate the equation to ensure all scale parameters, $\sigma_{i,t}$,
are positive. The $\gamma$'s are estimated coefficients, three of which
are allowed to vary by strategy, constrained by normal distributions:

\begin{align}
\gamma_{0,j}  &\sim \mathrm{N} \left(\eta_0 + \eta_1 M_j + \eta_2 E_j, \sigma^2_{\gamma_0}\right),\\
\gamma_{1,j}  &\sim \mathrm{N} \left(\mu_{\gamma_1}, \sigma^2_{\gamma_1}\right),\\
\gamma_{2,j}  &\sim \mathrm{N} \left(\mu_{\gamma_2}, \sigma^2_{\gamma_2}\right).
\end{align}

\noindent
The mean of the $\gamma_{0,j}$ coefficients is
modelled with
two strategy-level predictors.
The symbols $M_j$ and $E_j$ represent the raw mean species diversity and
mean combined fishing season length (effort) for strategy $j$.
The coefficients $\eta_1$ and $\eta_2$ represent predictors
in this strategy-level regression
estimating effects across strategies.

We fit our models with Stan 2.11.1
[@hoffman2014; @standevelopmentteam2015; @standevelopmentteam2015a; @carpenter2015]
and R 3.3.1 [@rcoreteam2016].
Stan implements the
No-U-Turn Hamiltonian Markov chain Monte Carlo algorithm [@hoffman2014]
to perform Bayesian statistical inference.
We assigned weakly informative priors on all parameters:
$N(0, 2^2)$ priors on all 
$\beta$, $\gamma$, and $\eta$ parameters,
and half-$t(3, 0, 2)$ priors 
(i.e. degrees of freedom of 3 and scale of 2)
on all $\sigma$ parameters.
We ran four chains and 2,000 iterations and discarded
the first 1,000 iterations of each chain
as warm up (i.e. 4,000 final samples).
We checked for chain convergence visually with trace plots.
We also ensured that $\hat{R}<1.05$ (the potential scale reduction factor)
and that the effective sample size,
as calculated by the rstan package, 
was greater than 200 for all parameters [@gelman2014].
We verified that our model returned sensible estimates by
comparing our estimates to an ad-hoc two-stage
model fit with the lme4 R package [@bates2015],
and by fitting our model to simulated data
and recovering the true values (Supporting Information).
Code to reproduce our results (with a simulated data set)
is available at <https://github.com/NCEAS/pfx-commercial>.

# Results

### Fig. \ref{fig:ts} - limited entry + movement towards salmon and specialization in Alaska

- move towards specialization, salmon fig 1D (overall declining trend disappears if you exclude salmon)
- halibut most often combined with other permits (exclude permit holders that hold any halibut and about ninety percent of the remaining permit holders hold only one permit) fig 1c
- Salmon the most often on its own (exclude salmon and the proportioned single permit decreases to about 70%) fig 1c
- majority are single permit holders (70-90%) fig 1C, but the single permit holders make up only about half of the overall revenue (and surpassed half for the first time in 2013 and stayed that way) fig 1B

### Something about overall model fit 

- Parameters from hierarchical model (main: Fig. S\ref{fig:fe}, varying: S\ref{fig:re})
- Residuals clearly show pattern that we see in the parameter estimates (Fig. S\ref{fig:resid})
- Downside residuals look very similar and justify using the full data set
(upside and downside) (Fig. S\ref{fig:resid-downside})
- Mention the year-permit varying intercepts and what they mean 
(Fig. S\ref{fig:str-yrs}) --- good and bad years where 
everyone in that strategy tended to do better or worse ---
this accounts for that and therefore compares everything to 
the expected revenue for that year 

### Fig. \ref{fig:spokes} - add permits

- nearly universal reduction in variability (Fig. S\ref{fig:spoke-all})
- e.g., `r 100-spoke[[2]]`% [`r 100-spoke[[3]]`--`r 100-spoke[[1]]`% IQR] 
lower expected standard deviation
for halibut plus another permit vs. single halibut permits);
- usually associated with higher median revenue 
- but can be very expensive (provide some examples from website)

### Fig. \ref{fig:across} - across strategies

- clear exponential decline in standardized variability with strategies that are inherently more species diverse 
- not a clear effect of days fished at this level
- fishing permits with higher species diversity reduces variability
(`r h1[[2]]`% [`r h1[[1]]`--`r h1[[3]]`% 95% CI] reduction per unit
increase in mean effective species diversity);

### Fig. \ref{fig:within} - within strategies

- fish more days, make more money (Fig. S\ref{fig:fe})
- after accounting for days fished, most commonly, 
specializing is associated with higher revenue, 
but sometimes the reverse (Fig. \ref{fig:within})
- if we don't account for changes in effort, then higher species diversity 
means more money fairly consistently (TODO make figure)
- changing species diversity (number of species fished or evenness of species
fished) tends to be associated with greater variability / uncertainty
year-to-year (Figs. \ref{fig:within}, S\ref{fig:fe})
- increasing species diversity within
the constraints of a permit set has a fishery-dependent effect on revenue and
is usually (`r g2_dens`% probability)
associated with increased uncertainty in revenue.
(Fig. \ref{fig:within})

# Discussion

### Take homes / summary

- What's novel
    - Huge dataset
    - Permit level data
    - Account for covariates
    - Accounting for strategies
    - Keeping annual data (not condensing to mean and cv)
    - Joint modelling
    - look at different mechanisms of diversification
- Suggests it's mostly an across strategy/permit effect (lowering variability),
  which isn't always a possibility because of regulations and cost 
- Adjusting within current permit constraints is risky... Not clearly risk
  reducing to generalize from one year to the next
- not as simple as increase diversity (within permit) and reduce risk
- 2 dimensions to think about (revenue and variability); and the trade-off is fishery specific 
    - effects within a permit strategy are permit specific  
    - Importantly it's a tradeoff - 2 dimensions - sometimes make more money by specializing... Some highly variable single species permits make a ton of money... which preferable?

### Reconciling the effects at different levels of species diversity 

- How fit in with K and H?
- Year to year risky... But maybe evens out/better over time
- diversity within permit - often bycatch, limited room for intentional diversification
- Risk results not really risk, but unpredictability given our covariates ...
don't change anything and it's most predictable how well you'll do
- What's unpredictable in our model, might be more predictable on a year to
year, species by species basis for individual permit holder
- Equilibrium, average variability vs year to year effect for one person

### Limitations + obvious extensions

- limitations...
- Vary by time
- Vary by mean diversity
- Spatial variation
- Nonlinear effects
- Response to specific events like EVOS and ITQs

### Implications for limited entry, ITQs, and catch shares

- if this is true, this is concerning because in many areas of the world and
fisheries, regulations and market forces have encouraged specialization 
- Permit diversity result: a privileged position - expensive to get to into
(quote king crab license) 
- Locking into permits limits ability to pick lower risk strategies and is
getting people locked into low diversity permits that may be doing well now,
but not necessarily always

### Conclusions + recommendations

\bibliographystyle{apalike}
\bibliography{refs}

\clearpage

# Figures


\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.7\textwidth]{../figs/fig1-jk.pdf}
\caption{
Trends in revenue and diversification over time.
(A) Map of fishing regions in Alaska, USA.
(B) Total revenue in billions of 2009 U.S. dollars from `r range_yrs[1]` to
`r range_yrs[2]` for multiple (grey) and single (blue) permit holders.
(C) Proportion of single permit holders for all fisheries (thick black line)
and with permit holders omitted each year who fished any of the
`r n_jacknifed` top species groups (thin lines).
(D) Mean effective diversity of species fished (Simpson's diversity index)
for all fisheries (thick black line)
and with permit holders omitted each year who fished any of the
`r n_jacknifed` top species groups (thin lines).
}
\label{fig:ts}
\end{center}
\end{figure}

\clearpage

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.9\textwidth]{../figs/stan-gg-spoke2.pdf}
\caption{Diversifying permits reduces risk and usually increases revenue.
Two examples of expected revenue variability
and median revenue for
owners of (A) individual halibut and (B) gillnet herring roe permits
(dark blue)
compared with owners of the same permit plus one more (medium blue)
or two more permits (light blue).
On the y-axis, dots and lines represent posterior median and 95\%
credible intervals of expected standard deviation.}
\label{fig:spokes}
\end{center}
\end{figure}

\clearpage

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.7\textwidth]{../figs/stan-across-strategy-variability2.pdf}
\caption{Individuals who fish
permits and permit combinations (permit strategies) with higher species diversity
have lower expected revenue variability.
On the y-axis, dots, thick lines, and thin lines represent posterior medians,
50\%, and 95\% credible intervals for
the `r nstr` most common permit strategies.
Thick black line, and dark and light grey shaded regions
indicate median, 50\%, and 95\% credible intervals
of strategy-level regression built into the hierarchical model.
Estimated variability represents expected standard deviation
for a permit holder who does not change
species diversity or days fished from year to year
with the strategy-level effect of mean days fished removed.
}
\label{fig:across}
\end{center}
\end{figure}

\clearpage

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=1.1\textwidth]{../figs/stan-offset-break-anti-spaghetti.pdf}
\caption{Effects of (A) generalizing and (B) specializing from year to year for
individual permit holders across the `r nstr` most common strategies.
Changing species diversity from year to year
is associated with higher variability
for individuals within the strategies above the zero line.
Generalizing (A) or specializing (B) is associated with
greater expected revenue to the right of the zero line
and lower expected revenue to the left of the zero line.
These effects are estimated holding
fishing season length constant from year to year.
Dots and line segments represent medians and
50\% credible intervals of posteriors.
}
\label{fig:within}
\end{center}
\end{figure}

\clearpage

# Notes

## @kasperski2013
- vessel-level CFEC data
- dome-shaped result
- Herfindahl-Hirschman Index/HHI (Simpson)
- diversity decreasing for those still fishing in 2010, but not otherwise

## @sethi2012b
- risk measures to 90 Alaskan fishing fishery/fleets
- revenue more variable than catch (+ price = extra variability)
- large catches, big fleets = lower risk

## @sethi2010b
- risk management for fisheries review

## @sethi2012
- equal-share catch-pooling cooperatives in king and snow crab
- "variability reduction depends on the degree of correlation amongst
  participants’ catches"

## @sethi2014
- community-level CFEC data
- number of fisheries strongest predictor of (lower) CV/CVaR
- also simpson index
- suggest barriers to commercial entry increase community risk by restricting
  access to 'variety of fisheries'
- some communities have limited ability to diversify; may need to diversify
  outside of fishing

[@perruso2005]
[@hilborn2001]
[@edwards2004]
[@trenkel2013]
[@mumford2009]
[@oostenbrugge2002]
[@baldursson1997]
[@jardine2015]
[@sanchirico2008]
