```{r set-knitr-options, echo=FALSE}
library("knitr")
opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, results="hide")
```
```{r get-dat}
load("../data-generated/output-values.rda")
h1 <- readRDS("../data-generated/h1_effect.rds")
spoke <- readRDS("../data-generated/spoke_fraction.rds")
g2_dens <- readRDS("../data-generated/g2_prob_dens_above_zero.rds")
n_grouped_permits <- readRDS("../data-generated/n_grouped_permits.rds")
```

<!-- Diversified portfolios of fishing permits and species reduce risk for fishermen, -->
<!-- but changing species diversity can be risky, costly, and limited by regulations. -->
<!-- Goal: Introduction, Methods, Results, Discussion, Conclusion:
<!-- 600, 900, 500, 800, 200, -->

# Abstract

<!-- TODO slash length of abstract -->

A fundamental challenge for individuals who rely on the extraction of natural
resources for their livelihood is income variability. Fishers, loggers, farmers,
and miners, among others, face high levels of income uncertainty driven by a mix
of environmental, biological, and economic variability. A key component
underlying the sustainability of these industries is identifying both how
regulatory actions and individual behaviour affect income variability and by
extension the economic stability and sustainable use of natural resources. For
commercial fisheries, recent work has identified that targeting a diversity of
species can reduce income variability. However, any diversity-variability
relationship is embedded in a regulatory framework that restricts individual
ability to change fishing behaviour. Here we identify three primary axes by
which fishers in Alaska, one of the world's most productive fishery regions, can
potentially diversify fishing activities and investigate how individual actions
along each diversity axis affect both revenue and revenue variability. We
construct a novel statistical model to simultaneously assess the value to
individual fishers of choosing to (1) own permits that allow targeting a diverse
fish species assemblage, (2) diversify by owning multiple permits, or (3) catch
a greater species diversity within the constraints of a given permit. We show
contrasting effects of diversity on revenue and revenue variability along the
three axes. First, we show that fishing permits with higher species diversity
reduces variability (`r h1[[2]]`% [`r h1[[1]]`--`r h1[[3]]`% 95% CI] reduction
per unit increase in mean effective species diversity). Next, we show that
fishing an additional permit is associated with higher revenue and lower
variability (e.g., `r 100-spoke[[2]]`% [`r 100-spoke[[3]]`--`r 100-spoke[[1]]`%
IQR] decline in standard deviation for fishers with halibut and another permit
vs. only a halibut permit). Finally, we show that increasing species diversity
within the constraints of a permit set has a fishery-dependent effect on revenue
and is usually (`r g2_dens`% probability) associated with increased uncertainty
in revenue the following year. Thus, while there are clear benefits to some axes
of diversification, others are less effective. Importantly, the most effective
mechanism for increasing revenue and decreasing risk --- purchasing additional
permits to participate in additional fisheries --- was also the most costly.
Furthermore, the least costly option of increasing catch diversity within an
existing permit was associated with increased revenue variability. Thus, over
the past 30 years as the Alaskan commercial fishing fleet has become
increasingly specialized and most individuals only participate under a single
permit, the practical benefits accrued from increased species diversification
are limited. 

Our results suggest that the benefit of species diversity for revenue
variability is mostly an across-permit effect, not a benefit that can be easily
harnessed by an individual permit holder. Effective methods of diversification
--- purchasing additional permits or switching permits --- are costly and often
limited by regulations such as limited entry programs and individual
transferable quotas (ITQs). Take home ...

\clearpage

# Introduction

<!-- Papers to cite [@smith1986] -->

Managing natural resources and sustaining the livelihoods of those that depend
on their extraction is hard. These livelihoods tend to have high year-to-year
variability in income relative to other professions [@mishra2002; @sethi2012b].
In addition to economic risk, such as changes in demand or prices, individuals
dependent on natural resources are also subject to biological and environmental
variability. For example, drought and flooding are a major source of risk for
agricultural food security and farmers' incomes [@gregory2005], and catastrophic
disease outbreaks and wildfires increase risk for the logging industry
[@kirilenko2007].

Individuals who rely on natural resources for income develop strategies to
reduce income variability. For example, farmers may rotate their crops to buffer
short-term climate variation and long-term environmental trends [@zentner1988].
Similarly, loggers maintain a diverse mix of tree species, age structure, and
harvest in geographical areas to buffer against disease outbreak, wildfires, and
price variation (REF). However, external forces may restrict the available
options to dampen income variability. Otherwise well-intentioned regulations may
limit how individuals diversify, or incentivize against diversification. For
example, in terrestrial agriculture, crop subsidies in the US may incentivize
some farms to specialize (e.g. dairy, grains, soybeans; @mishra2002). In other
instances, external market forces (e.g. prices or demand) may discourage the
adoption of diverse strategies and help promote specialization [@bradshaw2004]. 

Commercial fishers present an interesting example of diversification, given the
extreme risk that some individuals experience, particularly relative to
terrestrial agriculture [@sethi2012b]. The majority of risk insurance programs
available to terrestrial farmers (e.g. crop insurance programs) are not
obtainable by commercial fishers [@kasperski2013]. However, fishers maintain the
ability to lower risk by diversifying their fishing portfolio. They can, for
example, target as many species as permitted by a license, own multiple
licenses, or diversify spatially and fish in multiple regions. The effectiveness
of these diversification strategies for increasing revenue and reducing revenue
variability remains uncertain. For example, theory and empirical studies suggest
there can be substantial benefits to specialization for both farmers and fishers
[@abdulai2000; @arnason2005]. Furthermore, the majority of research examining
the effects of diversification on revenue do so at aggregate levels that combine
individuals (e.g. households, counties, or communities; [@kasperski2013;
@sethi2014]). Thus, key uncertainties remain about the fundamental relationship
between diversity, fishing revenue, and fishing revenue variability at the level
of the individual.

The state and federal fisheries of Alaska provide a unique opportunity to
understand the role of individuals' strategies in determining income and income
variability. Fisheries in Alaska are among the highest-valued fisheries in the
world with ~2.7 billion kg landed biomass valued at nearly 2 billion USD
annually [@nmfs2016]. Like much of the rest of the world, the days of open,
unregulated fisheries in most developed countries have passed --- fishers must
own a license to participate in virtually all fisheries [@hilborn2003a]. And
while such regulations have helped reduce overfishing, they limit the available
strategies to reduce income variability. For example, many permits limit fishing
to a single species, diversifying by purchasing additional fishing permits can
be expensive (e.g. ~\$192K USD for a king and tanner crab permit in Southeast
Alaska in 2016; <https://www.cfec.state.ak.us/pmtvalue/X_K69A.HTM>), and some
fisheries limit the entry of new participants through individual fishing quotas
(IFQs) [@grimm2012]. 

Using a unique dataset collected over 30 years (1985--2014) on individuals,
rather than fleets or vessels, we examine how revenue and revenue variability of
individuals is affected by diversification. By tracking individuals we can track
people that switch fishery permits over time or use multiple vessels. Similarly,
we can disentangle fishers that share the same vessel. We develop and apply
a hierarchical Bayesian model to understand the effect of fished species
diversity [OLE: Call this diversity of catch rather than fished?] on individual 
revenue and revenue variability that builds on results
from vessel [@kasperski2013], fleet [@sethi2012b], and community-level
[@sethi2014] variability. 

We develop a variance-function regression model [@davidian1987; @hall1989], to
simultaneously models changes in the mean and variance of individual revenue as
a function of covariates. We simultaneously estimate the effect of species
diversity (the inverse of Simpson's diversity index [@simpson1949; @jost2006])
on individual revenue and revenue variability, and estimate effects both within
and across fishing strategies while controlling for fishing effort and latent
time effects (e.g. changing environmental conditions). To our knowledge, this is
the first application of this type of model in any natural resource field.

# Results

We evaluated trends in revenue and diversification over time for individuals
fishing commercially in Alaska from 1985--2014 (Fig. \ref{fig:ts}A). The
majority (70--90%) of fishers hold single permits, with a slight increase in
single-permit holders since 2000 (Fig. \ref{fig:ts}C). This overall trend is
most sensitive to halibut or salmon permit holders --- compared to other
fisheries, halibut permit holders are more likely to hold halibut plus other
permits, whereas salmon permit holders are more likely to hold only one permit
(Fig. \ref{fig:ts}C). Despite making up only 10--30% of the individuals,
multiple permit holders earn approximately half the overall revenue (Fig.
\ref{fig:ts}B). The mean effective fished species diversity of this revenue
declined somewhat from 1980--2000 but has since remained steady (Fig.
\ref{fig:ts}D). However, this trend appears to be driven by an increase in
salmon specialists in the 1980s and 1990s --- the trend disappears if salmon
license holders are excluded (Fig. \ref{fig:ts}D).

Individual revenue variability decreases exponentially with higher mean fished
species diversity across permit sets (Fig. \ref{fig:across}). Controlling for
effort, strategies with one unit higher mean effective species diversity had on
average `r h1[[2]]`% (`r h1[[1]]`--`r h1[[3]]`% 95% CI) lower variability. This
pattern looks qualitatively similar with or without individuals targeting
salmon. As examples, variability was highest for the herring roe gillnet and
king crab fisheries. Most of the strategies with the lowest estimated revenue
variability combined halibut with a salmon permit and had high levels of fished
species diversity (Fig. \ref{fig:across}).

Comparing single permit holders to individuals holding that same permit plus
additional permits illustrates the expected reduction in revenue variability
from purchasing additional permits (Fig. \ref{fig:spokes}, Fig.
S\ref{fig:spoke-all}). For example, individuals fishing a halibut longline
permit plus one additional permit always had lower expected revenue variability
compared to individuals fishing only a halibut longline permit (median reduction
of `r 100-spoke[[2]]`% [`r 100-spoke[[3]]`--`r 100-spoke[[1]]`% IQR], Fig.
\ref{fig:spokes}A). Likewise, adding a permit to a herring roe gillnet strategy
substantially reduced expected revenue variability (Fig. \ref{fig:spokes}B).
Adding permits also usually increased expected revenue, although the permits
themselves represent a significant expense (e.g. ~\$192K USD for a king and
tanner crab permit in Southeast Alaska in 2016;
<https://www.cfec.state.ak.us/pmtvalue/X_K69A.HTM>).

In contrast to the clear across-strategy effects of species diversity on revenue
and revenue variability, the effect of changing fished species diversity for an
individual *within* the constraints of a given permit set is more variable (Fig.
\ref{fig:within}). After accounting for fishing effort, specializing tends to be
associated with increased revenue (Fig. \ref{fig:within}; Fig. S\ref{fig:fe},
Fig. S\ref{fig:re}) but also increased variability. On average, there is a ~0.5%
increase in the year-to-year revenue ratio per 1% decrease in species diversity,
and across strategies, `r g2_dens`% of the probability density supports
a positive relationship between specialization and revenue. 

However, adjusting (increasing or decreasing) an individual's fished species
diversity within the constraints of a permit set tends to be associated with
greater expected variability or uncertainty in year-to-year revenue (Fig.
\ref{fig:within} strategies above 0; Figs. S\ref{fig:fe}, \ref{fig:re},
\ref{fig:mirror}). This effect was strongest for individuals with a halibut
permit or halibut and sablefish permits. Individuals with these strategies
experienced a ~2% increase in variability for a 1% change in the year-to-year
ratio of species diversity (Fig. \ref{fig:within}). Becoming more generalized in
species diversity also leads to increased revenue variability (Fig.
S\ref{fig:mirror}A), suggesting that the act of changing strategies is
a contributor to increased variability of revenue. This benefit of specializing
is most apparent after accounting for fishing effort (Fig. S\ref{fig:noeffort}).

Alaskan halibut and sablefish fisheries were converted to individual fishery
quota (IFQ) management in 1994, which potentially confounds our estimates of the
effect of diversity and a major regulatory change. We refit our models allowing
all strategy-level effects to vary separately before and after the introduction
of IFQs for strategies involving halibut and sablefish permits. The introduction
of IFQs reduced variability by ~25% (XX--XX 95% CI) for individuals with only
halibut or sablefish permits (Fig. S\ref{fig:ifq}A). Furthermore, for most
strategies, the association between changing species diversity and increased
revenue variability became stronger after the introduction of IFQs (Fig.
S\ref{fig:ifq}B). This was especially true for individuals fishing only halibut
or sablefish and less so for individuals combining halibut or sablefish with
a salmon permit (Fig. S\ref{fig:ifq}B).

# Discussion

For individuals deriving livelihoods from natural resources, theory suggests
that adopting strategies with a diversity of assets provide substantial benefits
over strategies with lower diversity [REFS]. However, the ability to effectively
diversify in practice is limited by regulation and market forces. Here we
document the effect of changing species diversity on revenue and revenue
variability across three scale. Each scale reflects individual-level choices
that produce varied levels of species diversity. We find that individuals
fishing permit sets that allow high levels of species diversity tend to have
lower revenue variability. Adding an additional fisheries permit usually
increases revenue and decreases revenue variability. However, within the
constraints of a permit and accounting for effort, specializing (reducing
species diversity) tends to increase revenue and specializing or generalizing
tends to be associated with greater revenue variability, with considerable
differences across fisheries. Thus, the most effective mechanism by which
species diversity can lower revenue variability for individual fishers ---
switching or adding fishing permits --- is often costly, limited by regulations,
and consequently unavailable to many individuals in practice.

<!-- TODO mention the quadrants in figure 4 --> 

Our results shows additional support for the idea that individual fishers can
reduce their income variability by diversifying their catch portfolios. However,
most individuals do not appear to change their catch diversity substantially
from year to year, and most strategies adopted by Alaskan fishermen lie in
a high-risk, high-reward quadrant in which reducing diversification increases
both revenue variability and expected revenue. Previous work has suggested that
changes to fisheries management such as limited entry or catch share programs
may inadvertently act as barrier to diversification and promote specializing
(Hilborn et al. 2003; Kasperski and Holland 2013). While some of these
specialized fishing strategies may be associated with increased revenue
variability, the benefit may be that they are also associated with increased
revenue. External market forces may interact in concert with management
regulations to promote specialization for fishers or other individuals dependent
on natural resources. In terrestrial agriculture for instance, there are also
many instances of specializing being profitable; drivers include the ability to
specialize on high-profit cash crops (tobacco, cocoa), the availability of
government agricultural subsidies (soybeans, corn in the USA), and farming on
a large industrial scale. With increasing environmental variability and extreme
events in a changing climate (REF), it is particularly important that
individuals who rely on natural resources as a primary source of income have
strategies to reduce risk. Though variability in individual revenues may not be
actively incorporated into many management decisions, analyses like ours help to
understand the range of diversification strategies available and how their
benefits may change over time.

Our work compliments previous results that have examined the effects of species diversity at the 
community (Ref) and vessel(Ref) level that show THIS (Sethi, Kasperski, others)

### Reconciling the effects at different levels of species diversity

- why variability/uncertainty might increase with changes in diversity...

- and how fit in with @kasperski2013, @sethi2012b, @sethi2014

- part of the difference is individuals vs vessels. E.g. @kasperski2013
  suggested that diversity declined following IFQs for halibut sablefish
  vessels. The reason for the decline was that the vessels were allowed to fish
  halibut for more days (because people were sharing boats). At the level of
  people, we wouldn't expect the diversity of someone who only had
  halibut/sablefish permit to be affected by IFQs. The person, not the vessel
  holds the quota share: https://alaskafisheries.noaa.gov/fisheries/ifq

- year to year risky... But maybe evens out/better over time

- diversity within permit - often bycatch, limited room for intentional
  diversification (unless you have a M or salmon permit)

- risk results not really risk, but unpredictability given our covariates ...
  don't change anything and it's most predictable how well you'll do. What's
  unpredictable in our model, might be more predictable on a year to year,
  species by species basis for individual permit holder

- CV vs year-to-year effect for one person

Other:

- specializing can be profitable within some fisheries. Eric: analog here in
  farming literature is that diversification for small scale farmers generally
  good, except when they specialize on high demand/high value crops (tobacco,
  cocoa). Seems like benefits of diversification are also smaller for
  large-scale industrial farmers.


### Limitations + obvious extensions

There are a number of limitations to our model and extensions that could further
our understanding of the role of diversification...

- could vary by time and vary by mean diversity

- spatial/geographic variation

- nonlinear effects

- response to specific events like EVOS

### Implications for limited entry, ITQs, and catch shares

- concerning because in many areas of the world and fisheries, regulations and
  market forces have encouraged specialization / limited (effective)
  diversification (e.g. multiple permits or diverse permits)

- locking into permits limits ability to pick lower risk strategies and is
  getting people locked into low diversity permits that may be doing well now,
  but not necessarily always

### Conclusions + recommendations

- bring it out to wider than fisheries (farming, timber industry, ...)

- 2 dimensions to think about (revenue and variability); and the trade-off is
  fishery specific: effects within a permit strategy are permit specific;
  sometimes make more money by specializing... some highly variable single
  species permits make a ton of money

- broadly, diversification can provide a powerful form of protection from
  multiple sources of risk to income for those that rely on extracting natural
  resources, but not all forms of diversification are equal, and on an
  individual level there can be economic benefits to specializing and risks to
  adjusting ones strategy 

- with increasing environmental variability and extremes due to climate change
  those that rely on natural resources for their livelihoods need approaches to
  buffer risk more than ever 

- regulations that limit the ability to diversify can have other considerable
  benefits but it's important to be aware of trade-offs inherent in these
  regulations when making management decisions 

We fit a hierarchical variance-function regression model to 30 years of
commercial fisheries revenue data in Alaska to simultaneously estimate the
effect of changes in species diversity on individual's revenue and revenue
variability. 

\bibliographystyle{pnas2011}
\bibliography{refs}

\clearpage

# Figures

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.7\textwidth]{../figs/fig1-jk.pdf}

\caption{Trends in revenue and diversification over time. (A) Map of fishing
regions in Alaska, USA. (B) Total revenue in billions of 2009 U.S. dollars from
`r range_yrs[1]` to `r range_yrs[2]` for multiple (grey) and single (blue)
permit holders. (C) Proportion of single permit holders for all fisheries (thick
black line) and with permit holders omitted each year who fished any of the `r 
n_jacknifed` top species groups (thin lines). (D) Mean effective diversity of
species fished (Simpson's diversity index) for all fisheries (thick black line)
and with permit holders omitted each year who fished any of the `r n_jacknifed`
top species groups (thin lines). } \label{fig:ts} \end{center} \end{figure}

\clearpage

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.7\textwidth]{../figs/stan-across-strategy-variability2.pdf}

\caption{Individuals who fish permits and permit combinations (permit
strategies) with higher species diversity have lower expected revenue
variability. On the y-axis, dots, thick lines, and thin lines represent
posterior medians, 50\%, and 95\% credible intervals for the `r nstr` most
common permit strategies. Thick black line, and dark and light grey shaded
regions indicate median, 50\%, and 95\% credible intervals of strategy-level
regression built into the hierarchical model. Estimated variability represents
expected standard deviation for a permit holder who does not change species
diversity or days fished from year-to-year. }

\label{fig:across}
\end{center}
\end{figure}


\clearpage

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.9\textwidth]{../figs/stan-gg-spoke2.pdf}

\caption{Diversifying permits reduces risk and usually increases revenue. Two
examples of expected revenue variability and median revenue for owners of (A)
individual halibut and (B) gillnet herring roe permits (dark blue) compared with
owners of the same permit plus one more (medium blue) or two more (light
blue) permits. On the y-axis, dots and lines represent posterior median and 95\%
credible intervals of expected standard deviations of year-to-year revenue rates
of return.}

\label{fig:spokes}
\end{center}
\end{figure}

\clearpage

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.6\textwidth]{../figs/stan-offset-break-anti-spaghetti-specializing.pdf}

\caption{Effect of specializing on year-to-year changes in revenue and revenue
variability for individual permit holders \emph{within} the `r nstr` most common
strategies. Changing species diversity from year-to-year is associated with
higher variability for individuals within the strategies above the zero line.
Specializing is associated with greater expected revenue ratios to the right of
the zero line and lower expected revenue ratios to the left of the zero line.
For example, individuals with setnet permits for salmon in Prince William Sound,
Cook Inlet, Kodiak, Bristol Bay, or the Alaska Peninsula (second purple dot from
right) target primarily sockeye salmon and experience a $\sim$1.5\% increase in
their year-to-year revenue ratio (horizontal axis) and a $\sim$0.5\% increase in
their revenue variability (vertical axis) for a 1\% decrease in their ratio of
year-to-year species diversity. These effects are estimated holding fishing
season length constant from year-to-year. Dots and line segments represent
medians and 50\% credible intervals of posteriors. The effect of generalizing
(increasing species diversity) is similar (Fig. S\ref{fig:mirror}).}

\label{fig:within}
\end{center}
\end{figure}

\clearpage

# Materials and methods

<!-- TODO: get fig:re in results -->
<!-- Our model estimates overall average effects (Fig. S\ref{fig:fe}) and effects -->
<!-- that vary by permit set (Fig. S\ref{fig:re}). -->

We obtained commercial fisheries landings and revenue data for all permit
holders in Alaska from `r range_yrs[1]` to `r range_yrs[2]` from the Alaskan
Commercial Fisheries Entry Commission (CFEC). These data represent a total of
nearly `r round(ndat$initial_tot_rev/1e9)` billion 2009 USD and a combined `r
round(ndat$initial/1e6,1)` million rows of data. We adjusted for inflation by
setting all revenues to 2009 USD [@bea2016]. We implemented a number of
filtering steps to focus on individuals actively engaged in commercial fishing
(Supporting Information). In particular, we removed permit holders whose median
revenue was less than \$10,000 2009 USD to focus on individuals for whom
commercial fishing formed a substantial part of their income.

We used combinations of state and federal fishing permits held by an individual
in in a specific year to define fishing "strategies", representing the species
targeted (Supporting Information). For example, we combined permits that were
otherwise the same except for vessel size (e.g. a longliner fishing sablefish
on a vessel under 60 ft was considered to have the same strategy as a longliner
fishing sablefish on a vessel over 60 ft), and we combined a number of permits
that were targeting a single species and differed only in region fished.
Therefore, for example, someone fishing herring roe in Kodiak and someone
fishing herring roe in Alaska Peninsula were both considered to have a "herring
roe" permit group. For salmon permits, we formed salmon strategies based on
similarities in gear type and species compositon (Supporting Information).

After this process of combining permits, we were left with `r n_grouped_permits`
unique fishing permit groups (Table S\ref{tab:strategies}), which we aggregated
within each person-year combination to form permit strategies. For example, if
someone fished halibut and sablefish in the same year, their strategy for that
year would be "halibut-sablefish". Specific details on our strategy definitions
are available in the Supporting Information.

Our models focused on species diversity as a possible predictor of revenue and
revenue variability while controlling for effort. For consistency with previous
analyses [@kasperski2013], we calculated effective species diversity for
permit holder $i$ and time (year) $t$ as the inverse of Simpson's diversity
index $\lambda$ [@simpson1949; @jost2006] (also referred to in economics as the
Herfindahl-Hirschman Index) weighted by revenue $R$:

\begin{equation}
1/\lambda_{i,t} = 1 / \left(\sum\limits_{s=1}^{n_{i,t}}
  \left(R_{i,s,t}/ \sum\limits_{s=1}^{n_{i,t}} R_{i,s,t}\right)\right)^2,
\end{equation}

\noindent where $s$ indexes species from $1$ through $n$. We calculated effort
or effective season length as the sum of the season lengths for each permit
each year. Season length for a given permit was calculated as the number of
days between the first instance of fishing a permit and the last instance of
fishing a permit each year. For example, if someone fished for 30 days on a
salmon drift gillnet permit in Southeast Alaska and 40 days on a statewide
longline halibut permit, that individual's season length was calculated as 70
days.

## Hierarchical model

To jointly model revenue and revenue variability as a function of
individual-level covariates, we extended the basic variance-function regression
to a hierarchical Bayesian variance-function regression model with Gaussian
errors predicting $\log$ revenue, $\log(R_{i,t})$, for each fisher $i$ and time
$t$:

\begin{equation}
\log(R_{i,t}) \sim \mathrm{N} \left(\mu_{i,t}, \sigma_{i,t}^2 \right).
\end{equation}

\noindent Each year $t$ of revenue for fisher $i$ is assigned a strategy (permit
set) indexed by $j$ (Supporting Information, Table \ref{tab:strategies}) and
$\mu_{i,t}$ is modelled as:

\begin{equation}
\label{eq:mean}
\mu_{i,t} = \beta_{0,j,t} -
  \beta_{1,j} S_{i,t} I_{i,t} +
  \beta_{2,j} S_{i,t} I_{i,t} +
  \beta_3 D_{i,t} -
  \beta_4 S_{i,t} D_{i,t} I_{i,t} +
  \beta_5 S_{i,t} D_{i,t} I_{i,t} +
  \log(R_{i,t-1}),
\end{equation}

\noindent where $S_{i,t}$ is the log ratio of species diversity from
year-to-year, $\log \left(\mathrm{div}_{i,t}/\mathrm{div}_{i,t-1}\right)$,
$D_{i,t}$ is the log ratio of days fished from year-to-year, $\log
\left(\mathrm{days}_{i,t}/\mathrm{days}_{i,t-1}\right)$, and $I_{i,t}$ is an
indicator variable that takes a value $0$ if $\mathrm{div}_{i,t} \ge
\mathrm{div}_{i,t-1}$ (species diversity increasing; generalizing) and takes
a value $1$ if $\mathrm{div}_{i,t} < \mathrm{div}_{i,t-1}$ (species diversity
decreasing; specializing). Estimating a separate coefficient for the effect of
increasing diversity and decreasing diversity allows the costs and benefits of
these changes to be asymmetric.

We use $\log(R_{i,t-1}$) as an offset, and thus our model describes the log
ratio of revenue from year-to-year, i.e., $\log(R_{i,t}) - \log(R_{i,t-1})
= \log(R_{i,t}/R_{i,t-1})$. We model change in revenue because preliminary
analysis revealed non-stationary trends in revenue (resulting from experience or
other unmodeled factors). This approach eliminates having to deal with the
non-stationary time series and eliminates having to estimate thousands of
mean-revenue intercepts associated with individual fishers. We removed
person-year combinations without a subsequent year of revenue.

The $\beta$ terms represent estimated coefficients, some of which are allowed to
vary as "random effect" terms. The intercepts $\beta_{0,j,t}$ are allowed to
vary by strategy-year combinations, are centred on zero, and are constrained by
a normal distribution:

\begin{equation}
\beta_{0,j,t} \sim \mathrm{N} \left(0, \sigma^2_{\beta_{j,t}} \right).
\end{equation}

\noindent This approach, combined with the lack of a global intercept,
constrains the mean change in log revenue across all fishers and years to be
zero. I.e., we assume that the data are stationary after accounting for all
covariates and including an intercept that varies with strategy-year
combinations. Initial versions of the model confirmed that the global intercept
along with strategy-level intercepts would be estimated at almost exactly zero.
The estimated $\beta_{j,t}$'s represent years when people with a particular
permit set tended to collectively do better or worse than the long-term average
(Fig. S\ref{fig:str-yrs}). Therefore, an individual's expected change in revenue
(and variability) is after accounting for the general trend in revenue for that
permit set.

The slopes $\beta_{1,j}$ and $\beta_{1,j}$ are then allowed to vary by strategy
with estimated means:

\begin{align}
\beta_{1,j} &\sim \mathrm{N} \left(\mu_{\beta_1}, \tau^2_{\beta_1}\right),\\
\beta_{2,j} &\sim \mathrm{N} \left(\mu_{\beta_2}, \tau^2_{\beta_2}\right).
\end{align}

\noindent The $\beta_{1,j}$'s and $\beta_{2,j}$'s represent how much a 1\%
decrease and 1\% increase, respectively, in the ratio of species diversity from
year-to-year within a strategy translates to a given percent change in the ratio
of revenue from year-to-year. These are after controlling for changes in effort,
$\beta_3$, and the interaction between changes in effort and changes in species
diversity, $\beta_4$ and $\beta_5$.

Whereas traditional hierarchical linear models assume homoscedasticity --- that
the residual variance does not vary systematically --- our variance-function
regression model allows the residual variance to be heteroscedastic and vary as
a function of predictors. Instead of estimating a single residual error scale
term, $\sigma$, we model the scale of the residual error with another
hierarchical model with a similar form:

\begin{equation}
\label{eq:cv}
\sigma_{i,t} =
\exp \left( \gamma_{0,j} -
\gamma_{1,j} S_{i,t} I_{i,t} +
\gamma_{2,j} S_{i,t} I_{i,t} +
\gamma_3 D_{i,t} -
\gamma_4 S_{i,t} D_{i,t} I_{i,t} +
\gamma_5 S_{i,t} D_{i,t} I_{i,t}\right).
\end{equation}

\noindent Again, $S_{i,t}$ and $I_{i,t}$ represent, respectively, percent change
in species diversity and an indicator variable for increasing or decreasing
species diversity. We exponentiate the equation to ensure all scale parameters,
$\sigma_{i,t}$, are positive. Residuals from the mean component of our model
clearly show a "V"-shaped pattern that the variance component of our model can
capture (Fig. S\ref{fig:resid}). Furthermore, the downside residuals
(year-individual data points with lower than expected revenue) look
qualitatively similar to the complete set of residuals (upside and downside)
(Fig. S\ref{fig:resid-downside}). Therefore, it seemed appropriate to model the
variability as a proxy for revenue risk. 

The $\gamma$'s are estimated coefficients, three of which are allowed to vary by
strategy, constrained by normal distributions:

\begin{align}
\gamma_{0,j}  &\sim \mathrm{N} \left(\eta_0 + \eta_1 M_j + \eta_2 E_j, \tau^2_{\gamma_0}\right),\\
\gamma_{1,j}  &\sim \mathrm{N} \left(\mu_{\gamma_1}, \tau^2_{\gamma_1}\right),\\
\gamma_{2,j}  &\sim \mathrm{N} \left(\mu_{\gamma_2}, \tau^2_{\gamma_2}\right).
\end{align}

\noindent The mean of the $\gamma_{0,j}$ coefficients is modelled with two
strategy-level predictors. The symbols $M_j$ and $E_j$ represent the mean
species diversity and mean combined fishing season length (effort) for strategy
$j$. The coefficients $\eta_1$ and $\eta_2$ represent predictors in this
strategy-level regression estimating effects across strategies.

We fit our models with Stan 2.11.1 [@hoffman2014; @standevelopmentteam2015;
@standevelopmentteam2015a; @carpenter2015] and R 3.3.1 [@rcoreteam2016]. Stan
implements the No-U-Turn Hamiltonian Markov chain Monte Carlo algorithm
[@hoffman2014] to perform Bayesian statistical inference. We assigned weakly
informative priors on all parameters: $N(0, 2^2)$ priors on all $\beta$,
$\gamma$, and $\eta$ parameters, and half-$t(3, 0, 2)$ priors (i.e. degrees of
freedom of 3 and scale of 2) on all $\sigma$ and $\tau$. We ran four chains and
2,000 iterations and discarded the first 1,000 iterations of each chain as warm
up. We checked for chain convergence visually with trace plots, ensured that
$\hat{R}<1.05$ (the potential scale reduction factor) and that the effective
sample size, as calculated by the rstan R package [@standevelopmentteam2015],
was greater than 200 for all parameters [@gelman2014]. We verified that our
model returned sensible estimates by comparing our estimates to an ad-hoc
two-stage model fit with the lme4 R package [@bates2015], where we first fit
a mixed-effects regression model to the mean, and then modelled the absolute
residuals from that first model in a second mixed-effects model representing the
variance component. We also validated the model with simulated data, and were
able to recover the true values (Supporting Information). Code used in our
analysis is available at <https://github.com/NCEAS/pfx-commercial/portfolios/>.

<!-- # Notes -->

<!-- ## @kasperski2013 -->
<!-- - vessel-level CFEC data -->
<!-- - dome-shaped result -->
<!-- - Herfindahl-Hirschman Index/HHI (Simpson) -->
<!-- - diversity decreasing for those still fishing in 2010, but not otherwise -->

<!-- ## @sethi2012b -->
<!-- - risk measures to 90 Alaskan fishing fishery/fleets -->
<!-- - revenue more variable than catch (+ price = extra variability) -->
<!-- - large catches, big fleets = lower risk -->

<!-- ## @sethi2010b -->
<!-- - risk management for fisheries review -->

<!-- ## @sethi2012 -->
<!-- - equal-share catch-pooling cooperatives in king and snow crab -->
<!-- - "variability reduction depends on the degree of correlation amongst -->
<!--   participants’ catches" -->

<!-- ## @sethi2014 -->
<!-- - community-level CFEC data -->
<!-- - number of fisheries strongest predictor of (lower) CV/CVaR -->
<!-- - also simpson index -->
<!-- - suggest barriers to commercial entry increase community risk by restricting -->
<!--   access to 'variety of fisheries' -->
<!-- - some communities have limited ability to diversify; may need to diversify -->
<!--   outside of fishing -->

<!-- [@perruso2005] -->
<!-- [@hilborn2001] -->
<!-- [@edwards2004] -->
<!-- [@trenkel2013] -->
<!-- [@mumford2009] -->
<!-- [@oostenbrugge2002] -->
<!-- [@baldursson1997] -->
<!-- [@jardine2015] -->
<!-- [@sanchirico2008] -->

<!-- vim: textwidth=80 
-->
