```{r set-knitr-options, cache=FALSE, echo=FALSE}
library("knitr")
opts_chunk$set(cache=TRUE, echo=FALSE, message=FALSE, autodep=TRUE,
  results='hide')
opts_chunk$set(warning=FALSE, message=FALSE, cache.path="../cache/",
  cache.lazy=TRUE)
```
```{r get-dat, cache=FALSE}
load("../data-generated/output-values.rda")
```

\noindent
Main message:
Diversified portfolios of fishing permits and species reduce risk for fishermen,
but changing species diversity can be risky, costly, and limited by regulations.

# Abstract

<!-- Fishing is inherently risky and we have long-sought regulations and fishing behaviors that reduce this risk. --> 
<!-- The increasing the diversity of species fished has been proposed to reduce risk for fishermen --> 

Individuals relying on natural resources as their primary source of income
often experience high levels of income variability 
due to the combined effects of environmental and price variation. 
Commercial fishers are a prime example and
recent work has sought to identify regulations and fishing behaviours 
that can reduce this variability. 
A top candidate has been that fishing higher levels of species diversity
decreases income variability.
However, it is unknown if this species-diversity effect 
on income variability holds 
across multiple approaches to diversifying catch,
at the level of individual permit holders,
and across different fishing strategies. 
Here, we simultaneously estimate the effects of three 
methods of diversification for Alaskan fishery permit holders
on revenue and revenue variability:
(1) owning multiple permits, 
(2) owning permits that inherently fish higher species diversity, 
and (3) adjusting species diversity within the constraints of a permit.
We control for changes in effort and other 
residual effects through time.
We find that 
fishing an additional permit is associated with 
lower variability (~XX--XX%) and higher revenue,
fishing permits with higher species diversity reduces risk 
(XX--XX% per unit increase in Simpson's diversity index),
but that changing species diversity within a
permit's constraints has a fishery-dependent effect on revenue and 
is usually (XX%) associated with increased revenue variability.
At the same time, Alaskan commercial fisheries
have increasingly focussed on 
single-permit (often salmon) fishing strategies
with single-permit fishers earning 
more than multiple-permit fishers
since 2013.
Our results suggest that the benefit of species diversity
on reducing revenue variability 
is mostly an across-permit effect, not
a benefit that can be easily harnessed by an individual permit holder.
Effective methods of diversification --- 
purchasing additional permits or switching permits ---
are costly and often limited by regulations such as limited entry
programs and individual transferable quotas (ITQs).
Take home ...

\bigskip
\noindent
Extra:

- there is often limited room for intentional diversification within a permit
- in fact, specializing can be profitable within some fisheries

<!-- , which is concerning  -->
<!-- given that some fishing regulations and market forces -->
<!-- may be increasingly promoting specialization. -->
<!-- Main message in 25 words or less: -->

<!-- Take 1: -->
<!-- - There is a trend towards specializing on single salmon permits in Alaska -->
<!-- - Diversified portfolios of fishing permits and species reduce risk, -->
<!--   but changing species diversity for an individual permit holder can be risky -->
<!-- - The types of diversity needed to reduce risk are often restricted by -->
<!--   cost and regulations -->

<!-- Take 2: -->
<!-- - The effect of species diversification on revenue and risk for fishermen -->
<!-- depends on the type of diversification and the fishery. -->

<!-- Introduction, Methods, Results, Discussion, Conclusion:
<!-- 600, 900, 500, 800, 200, --> 

<!-- write out each paragraph’s main message in 15 words or less --> 

<!-- circulate skeleton ... this is the time to fix problems with flow, -->
<!-- ideas/content, main message -->


# Introduction

What's been done

[@kasperski2013]
[@sethi2012b]
[@sethi2010b]
[@sethi2014]
[@perruso2005]
[@hilborn2001]
[@edwards2004]
[@trenkel2013]
[@mumford2009]
[@oostenbrugge2002]
[@baldursson1997]
[@jardine2015]
[@sanchirico2008]

Multiple ways one can diversify: 
buy another permit, pick diverse permit, diversify within existing constraints.

# Methods

[@standevelopmentteam2015]
[@rcoreteam2016]
[@simpson1949]

# Results

# Discussion

\bibliographystyle{apalike}
\bibliography{refs}

\clearpage

# Figures

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.7\textwidth]{../figs/fig1-jk.pdf}
\caption{
Trends in revenue and diversification over time.
(A) Map of fishing regions in Alaska, USA.
(B) Total revenue in billions of 2009 U.S. dollars from `r range_yrs[1]` to
`r range_yrs[2]` for multiple (grey) and single (blue) permit holders.
(C) Proportion of single permit holders for all fisheries (thick black line)
and with permit holders omitted each year who fished any of the
`r n_jacknifed` top species groups (thin lines).
(D) Mean effective diversity of species fished (Simpson's diversity index) 
for all fisheries (thick black line)
and with permit holders omitted each year who fished any of the
`r n_jacknifed` top species groups (thin lines).
}
\label{fig:ts}
\end{center}
\end{figure}

\clearpage

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.9\textwidth]{../figs/stan-gg-spoke2.pdf}
\caption{Diversifying permits reduces risk and usually increases revenue.
Two examples of expected revenue variability 
and median revenue for
owners of (A) individual halibut and (B) gillnet herring roe permits
(light blue)
compared with owners of the same permit plus one more (medium blue)
or two more permits (dark blue).
On the y-axis, dots and lines represent posterior median and 90\%
credible intervals.}
\label{fig:spokes}
\end{center}
\end{figure}

\clearpage

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.7\textwidth]{../figs/stan-across-strategy-variability.pdf}
\caption{Individuals who fish
permits and permit combinations (permit strategies) with higher species diversity
have lower expected revenue variability.
On the y-axis, dots, thick lines, and thin lines represent posterior medians,
50\%, and 95\% credible intervals for
the `r nstr` most common permit strategies.
Thick black line, and dark and light grey shaded regions
indicate median, 50\%, and 95\% credible intervals
of strategy-level regression built into the hierarchical model.
Estimated variability represents expected standard deviation
at an average of about
`r mean_permit_days` combined fishing days across permits,
and for a permit holder who does not change
species diversity or days fished from year to year.
}
\label{fig:across}
\end{center}
\end{figure}

\clearpage

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=1.1\textwidth]{../figs/stan-offset-break-anti-spaghetti.pdf}
\caption{Effects of (A) generalizing and (B) specializing from year to year for
individual permit holders across the `r nstr` most common strategies.
Changing species diversity from year to year 
is associated with higher variability
for individuals within the strategies above the zero line.
Generalizing (A) or specializing (B) is associated with
greater expected revenue to the right of the zero line
and lower expected revenue to the left of the zero line.
These effects are estimated holding
fishing season length constant from year to year.
Dots and line segments represent medians and 
50\% credible intervals of posteriors.
}
\label{fig:within}
\end{center}
\end{figure}

