```{r set-knitr-options, cache=FALSE, echo=FALSE}
library("knitr")
opts_chunk$set(cache=TRUE, echo=FALSE, message=FALSE, autodep=TRUE,
  results='hide')
opts_chunk$set(warning=FALSE, message=FALSE, cache.path="../cache/",
  cache.lazy=TRUE)
```


```{r get-dat, cache=FALSE}
load("../data-generated/output-values.rda")
```

# Introduction

[@kasperski2013]
[@simpson1949]
[@sethi2012b]
[@sethi2014]


# Methods

[@standevelopmentteam2015]

# Results

# Discussion

\bibliographystyle{apalike}
\bibliography{refs}

\clearpage

# Figures

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.7\textwidth]{../figs/fig1-jk.pdf}
\caption{Caption here}
\label{fig:ts}
\end{center}
\end{figure}

\clearpage

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.9\textwidth]{../figs/stan-gg-spoke2.pdf}
\caption{Caption here}
\label{fig:spokes}
\end{center}
\end{figure}

\clearpage

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.7\textwidth]{../figs/stan-across-strategy-variability.pdf}
\caption{Caption here}
\label{fig:across}
\end{center}
\end{figure}



\clearpage

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=1.1\textwidth]{../figs/stan-offset-break-anti-spaghetti.pdf}
\caption{Caption here}
\label{fig:within}
\end{center}
\end{figure}

