```{r set-knitr-options, echo=FALSE}
library("knitr")
opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, results="hide")
```
```{r get-dat}
load("../data-generated/output-values.rda")
h1 <- readRDS("../data-generated/h1_effect.rds")
spoke <- readRDS("../data-generated/spoke_fraction.rds")
g2_dens <- readRDS("../data-generated/g2_prob_dens_above_zero.rds")
n_grouped_permits <- readRDS("../data-generated/n_grouped_permits.rds")
```

<!-- Introduction, Methods, Results, Discussion, Conclusion -->
<!-- 600, 900, 500, 800, 200, -->

# Abstract

<!-- TODO < 250 words -->

Individuals who rely on the extraction of natural resources for their
livelihood --- fishers, farmers, miners, and loggers, among others --- face
high levels of income variability driven by a mix of environmental, biological,
and economic variability. A key component to managing these industries is
identifying how regulatory actions and individual behaviour affect income
variability and by extension the economic stability and sustainable use of
natural resources. For commercial fisheries, recent work has identified that
communities and vessels that fish a greater diversity of species have less
revenue variability than those that fish a lower diversity. However, it
is unclear if these benefits extend to the level of the individual and how
year-to-year changes in diversification affect revenue and revenue variability.
Here we identify three axes by which fishers in Alaska, one of the world's most
productive fishing regions, can diversify fishing activities.
We show that fishing permits with higher species diversity reduces individual
revenue variability and fishing an additional permit is associated with higher
revenue and lower variability. However, increasing species diversity within the
constraints of a permit set has a fishery-dependent effect on revenue and is
usually (`r g2_dens`% probability) associated with increased revenue
uncertainty the following year. Our results demonstrate that the most effective
option for individuals to decrease revenue variability is to purchase permits
to participate in additional or more diverse fisheries. However, this option is
costly and often limited by regulations such as limited entry and catch share
programs. With increasing environmental variability in a changing climate, it
will be particularly important that individuals who rely on natural resources
for their livelihood have effective strategies to reduce income variability.

<!-- TODO: fit something about specialization over time into the abstract.
Thus, over the past 30 years as the Alaskan commercial fishing fleet has become
increasingly specialized, -->

\clearpage

# Introduction

Managing natural resources and sustaining the livelihoods of those that depend
on their extraction is hard. These livelihoods tend to have high annual
variability in income relative to other professions [@mishra2002; @sethi2012b].
In addition to economic risk, such as changes in demand or prices, individuals
dependent on natural resources are also subject to biological and environmental
variability [@smith1986]. For example, drought and flooding are a major source
of risk for agricultural food security and farmers' incomes [@gregory2005], and
catastrophic disease outbreaks and wildfires increase risk for the logging
industry [@kirilenko2007].

Individuals who rely on natural resources for income develop strategies to
reduce income variability. For example, farmers may rotate their crops to buffer
short-term climate variation and long-term environmental trends [@zentner1988].
Similarly, loggers maintain a diverse mix of tree species, age structure, and
harvest in geographical areas to buffer against disease outbreak, wildfires, and
price variation (REF). However, external forces may restrict the available
options to dampen income variability. Otherwise well-intentioned regulations may
limit how individuals diversify, or incentivize against diversification. For
example, in terrestrial agriculture, crop subsidies in the U.S. may incentivize
some farms to specialize (e.g. dairy, grains, soybeans; @mishra2002). In other
instances, external market forces (e.g. prices or demand) may discourage the
adoption of diverse strategies and help promote specialization [@bradshaw2004].

Commercial fishers present an interesting example of diversification, given the
extreme risk that some individuals experience, particularly relative to
terrestrial agriculture [@sethi2012b]. The majority of risk insurance programs
available to terrestrial farmers (e.g. crop insurance programs) are not
obtainable by commercial fishers [@kasperski2013]. However, fishers maintain the
ability to lower risk by diversifying their fishing portfolio. They can, for
example, target as many species as permitted by a license, own multiple
licenses, or diversify spatially and fish in multiple regions. The effectiveness
of these diversification strategies for increasing revenue and reducing revenue
variability remains uncertain. For example, theory and empirical studies suggest
there can be substantial benefits to specialization for both farmers and fishers
[@smith1986; @abdulai2000; @arnason2005]. Furthermore, the majority of research
examining the effects of diversification on revenue do so at aggregate levels
that combine individuals (e.g. counties or communities;
[@kasperski2013; @sethi2014]). Thus, key uncertainties remain about the
fundamental relationship between diversity, fishing revenue, and fishing
revenue variability at the level of the individual.

The state and federal fisheries of Alaska provide a unique opportunity to
understand the role of individuals' strategies in determining income and income
variability. Fisheries in Alaska are among the highest-valued fisheries in the
world with ~2.7 billion kg landed biomass valued at nearly 2 billion USD
annually [@nmfs2016]. The days of open, unregulated fisheries in most developed
countries have passed --- fishers must own a license to participate in
virtually all fisheries [@hilborn2003a]. While such regulations have helped
reduce overfishing, they limit the available strategies to reduce income
variability. For example, many permits limit fishing to a single species,
diversifying by purchasing additional fishing permits can be expensive (e.g.
~\$192K USD for a king and tanner crab permit in Southeast Alaska in 2016;
<https://www.cfec.state.ak.us/pmtvalue/X_K69A.HTM>), and some fisheries limit
the entry of new participants through individual fishing quotas (IFQs)
[@grimm2012].

Using a unique dataset collected over 30 years on individuals,
rather than fleets or vessels, we examine how revenue and revenue variability of
individuals is affected by diversification. By tracking individuals we can track
people that switch fishery permits over time or use multiple vessels. Similarly,
we can disentangle fishers that share the same vessel. We develop and apply
a hierarchical Bayesian model to understand the effect of fisheries catch
species diversity on individual revenue and revenue variability that builds on
results from vessel [@kasperski2013], fleet [@sethi2012b], and community-level
[@sethi2014] variability.
We develop a variance-function regression model [@davidian1987; @hall1989], to
simultaneously models changes in the mean and variance of individual revenue as
a function of covariates. We simultaneously estimate the effect of species
diversity (the inverse of Simpson's diversity index [@simpson1949; @jost2006])
on individual revenue and revenue variability, and estimate effects both within
and across fishing strategies while controlling for fishing effort and latent
time effects (e.g. changing environmental conditions). To our knowledge, this is
the first application of this type of model in any natural resource field.

# Results

We evaluated trends in revenue and diversification over time for individuals
fishing commercially in Alaska from 1985--2014 (Fig. \ref{fig:ts}A). The
majority (70--90%) hold single permits, with a slight increase in single-permit
holders since 2000 (Fig. \ref{fig:ts}C). This overall trend is most sensitive to
halibut or salmon permit holders --- compared to other fisheries, halibut permit
holders are more likely to hold halibut plus other permits, whereas salmon
permit holders are more likely to hold a single permit (Fig. \ref{fig:ts}C).
Despite making up only 10--30% of the individuals, multiple permit holders earn
approximately half the overall revenue (Fig. \ref{fig:ts}B). The mean species
diversity of this revenue declined somewhat from 1980--2000 but has since
remained steady (Fig. \ref{fig:ts}D). However, this trend appears largely
driven by an increase in salmon specialists in the 1980s and 1990s --- the
trend disappears if salmon license holders are excluded (Fig. \ref{fig:ts}D).

Individual revenue variability decreases exponentially with higher mean fished
species diversity across permit sets (Fig. \ref{fig:across}). Controlling for
effort, strategies with one unit higher mean species diversity had on
average `r h1[[2]]`% (`r h1[[1]]`--`r h1[[3]]`% 95% CI) lower variability. This
pattern looks qualitatively similar with or without individuals targeting
salmon. As examples, variability was highest for the herring roe gillnet and
king crab fisheries. Most strategies with the lowest estimated revenue
variability combined halibut with a salmon permit and had high levels of fished
species diversity (Fig. \ref{fig:across}).

Comparing single permit holders to individuals holding that same permit plus
additional permits illustrates the expected reduction in revenue variability
from purchasing additional permits (Fig. \ref{fig:spokes}, Fig.
S\ref{fig:spoke-all}). For example, individuals fishing a halibut permit plus
one additional permit always had lower expected revenue variability compared to
individuals fishing only a halibut permit (median reduction of `r
100-spoke[[2]]`% [`r 100-spoke[[3]]`--`r 100-spoke[[1]]`% IQR], Fig.
\ref{fig:spokes}A). Likewise, adding a permit to a herring roe gillnet strategy
substantially reduced expected revenue variability (Fig. \ref{fig:spokes}B).
Adding permits also usually increased expected revenue, although the permits
themselves represent a significant expense (e.g. ~\$192K USD for a king and
tanner crab permit in Southeast Alaska in 2016
<https://www.cfec.state.ak.us/pmtvalue/X_K69A.HTM>).

In contrast to the clear across-strategy effects of species diversity on revenue
and revenue variability, the effect of changing fished species diversity for an
individual *within* the constraints of a given permit set is more variable (Fig.
\ref{fig:within}). After accounting for fishing effort, specializing tends to be
associated with increased revenue but also increased variability (Fig.
\ref{fig:within}; Figs S\ref{fig:fe}, S\ref{fig:re}, S\ref{fig:noeffort}). On
average, there is a ~0.5% increase in the year-to-year revenue ratio per 1%
decrease in species diversity, and across strategies weighted equally, `r g2_dens`% of the
probability density supports a positive relationship between specialization and
revenue.

Changing (increasing or decreasing) an individual's fished species
diversity within the constraints of a permit set tends to be associated with
greater expected variability or uncertainty in year-to-year revenue (Fig.
\ref{fig:within} strategies above zero; Figs. S\ref{fig:fe}, S\ref{fig:re},
S\ref{fig:mirror}). This effect was strongest for individuals with a halibut
permit or halibut and sablefish permits. Individuals with these strategies
experienced a ~2% increase in variability for a 1% change in the year-to-year
ratio of species diversity (Fig. \ref{fig:within}). Becoming more generalized in
species diversity also leads to increased revenue variability (Fig.
S\ref{fig:mirror}A), suggesting that the act of changing strategies is
a contributor to increased variability of revenue.

<!-- TODO mention the quadrants in figure 4 -->

Alaskan halibut and sablefish fisheries were converted to individual fishery
quota (IFQ) management in 1994, which potentially confounds our estimates with
a major regulatory change. Therefore, we refit our models allowing all
strategy-level effects involving halibut and sablefish permits
to vary separately before and after the introduction of IFQs. The introduction
of IFQs reduced variability by ~25% (TODO XX--XX 95% CI) for individuals with
only halibut or sablefish permits (Fig. S\ref{fig:ifq}A). Furthermore, for most
strategies, the association between changing species diversity and increased
revenue variability became stronger after the introduction of IFQs (Fig.
S\ref{fig:ifq}B). This was especially true for individuals fishing only halibut
or sablefish and less so for individuals combining halibut or sablefish with
a salmon permit (Fig. S\ref{fig:ifq}B).

# Discussion

For individuals deriving livelihoods from natural resources, theory suggests
that adopting strategies with a diversity of assets provides substantial
benefits over strategies with lower diversity [REFS]. However, in practice, the
ability of individuals to effectively diversify is limited by regulation and
market forces. Here we document the effect of individual-level changes in
species diversity on revenue and revenue variability across three scales. We
find that (1) individuals with fishing permit sets that allow high levels of
species diversity tend to have lower revenue variability. (2) Adding a fisheries
permit usually increases revenue and decreases revenue variability. However, (3)
within the constraints of a permit and accounting for effort, specializing tends
to increase revenue while both specializing and generalizing tend to correlate
with greater revenue variability with considerable differences amongst
fisheries. Thus, the most effective mechanism by which species diversity can
lower revenue variability for individual fishers --- switching or adding fishing
permits --- is costly, limited by regulations, and consequently unavailable to
many individuals.

Our results across fishing strategies are in line with previous findings at the
community and vessel levels [@kasperski2013; @sethi2012b; @sethi2014], but our
results with respect to individual year-to-year behaviour provide new insights.
Within strategies, specializing tended to be associated with increased revenue.
This could be for a number of reasons. First, specialization can increase
fishing efficiency [@smith1986]. Efficiency gains from specialization can be
seen in other natural resource sectors. For example, rice farmers in Ghana can
increase their efficiency by devoting more land and labor to rice production
compared to other crops [@abdulai2000]. Second, given the availability of
valuable resources, specializing can be highly profitable. For instance, when
cocoa prices have been high, Nigerian farmers have tended to specialize on the
crop [@mustapha1999] and specializing on valuable tobacco farming can be the
optimal strategy regardless of alternative crops [@chavez2014]. In our analyses,
fishers with the opportunity to fish highly valuable sockeye salmon benefited
the most from specializing. However, specializing can have longterm risks if
market, biological, or environmental conditions change. For example, U.S.
farmers specialized on wheat, corn, or soybeans were hit harder than diversified
farmers in the early 1980s when a financial crisis hit [@ikerd1990]. Third,
generalization --- a lack of specialization --- can happen as a consequence of
non-ideal circumstances such as poor availability of a more valuable fish stock
in a given year. For example, downward economic shocks have been associated with
increased diversity for households relying on forests for income [@debela2012].

We also found that altering species diversity tended to be associated with
increased revenue variability. There are number of possible reasons for this
relationship. For instance, changes to fishing patterns will sometimes be
associated with a learning or adaptation phase as individuals become accustomed
to new locations, new gear, or new species [REF]. Additionally, changes in
diversity might reflect years in which targeted species were less available and
fishers were forced to target other species or landed more bycatch adding
stochasticity to their revenue. Most importantly, this increased variability
reflects an inability for our large-scale model to predict an individual's
revenue in a subsequent year. There are many reasons, often unique to an
individual permit holder's behaviour [@hilborn1985] and therefore not included
in our large-scale model, that might make that specific individual's revenue
less predictable if species diversity changes.

There are many possible barriers to diversification for individuals relying on
natural resources for their livelihood. A primary barrier is the cost of
diversifying. For commercial fishers, purchasing additional licenses and gear
typically represents a substantial cost. Related barriers are apparent in other
natural resource sectors. For example, diversification has been positively
linked with total income or access to financial capital for farmers
[@abdulai2001; @barrett2001], small-scale fishing communities [@olale2012], and
those relying on forest resources [@illukpitiya2008]. Another major barrier to
diversification can be management restrictions. In commercial fisheries, while
catch shares [@costello2008] and limited season lengths [REF] can be effective
tools to avoid overfishing and distribute social capital, they can also limit
opportunities for diversification [@smith1986; @kasperski2013]. However, catch
shares, in particular, may also reduce income variability (Fig.
S\ref{fig:ifq}A) perhaps offsetting the effects of reduced diversification
[@holland2016].

Our analysis provides a synthetic view of the effects of species diversity on
individuals over 30 years in one of the most productive fishing regions in the
world. However, this large temporal and geographical scope potentially obscures
time or regional-variation in the consequences of diversification on revenue
and revenue variability. We expect future research examining diversification
over finer scales will uncover additional interesting relationships. Furthermore, in
addressing the overall pattern, we have averaged out several major discrete
events, such as the Exxon Valdez oil spill in 1989, that had significant
consequences for the ecological and social systems in Alaska [@paine1996]. In
addition to revenue or income variability, it is important to understand the
benefits of diversification in the context of rare and extreme disturbances
which occur in natural systems worldwide [REF].

<!--
Worldwide, fisheries confront the challenge of sustainable management (). While
many fisheries are well managed and productive, many are also rebuilding from
periods of over-exploitation (Worm et al. Science 2009). A common tactic to aid
in fisheries rebuilding is to modify regulations to reduce the unintended
harvest rebuilding stocks while allowing the continued harvest of healthy stocks
(need REF). Such bycatch reduction is an important goal for both ecological
(Crowder and Murawski 1998, Lewison et al. 2004) and economic (Curtis and Hicks
2000; Chakravorty and Nemoto 2001) reasons. Cumulatively, though, regulatory
changes for bycatch reduction as well as other reasons gradually lead to
fisheries that target one or a few species and individual fishers wishing to
participate in multiple fisheries must own (or purchase) multiple permits. Thus
a general consequence of fisheries regulations that are shown to improve
sustainability of fish stocks and avoid overfishing - including improved
assessments (Hilborn and Ovando 2014) and individual quota systems (Costello et
al. 2008 ) - may ultimately restrict individual diversification or at least make
developing a diverse fishing portfolio more costly. Add Alaska example of
IFQs...?
-->

<!-- Conclusions + recommendations

- bring it out to wider than fisheries (farming, timber industry, ...)
- 2 dimensions to think about (revenue and variability); and the trade-off is
  fishery specific: effects within a permit strategy are permit specific;
  sometimes make more money by specializing... some highly variable single
  species permits make a ton of money
- broadly, diversification can provide a powerful form of protection from
  multiple sources of risk to income for those that rely on extracting natural
  resources, but not all forms of diversification are equal, and on an
  individual level there can be economic benefits to specializing and risks to
  adjusting ones strategy
- with increasing environmental variability and extremes due to climate change
  those that rely on natural resources for their livelihoods need approaches to
  buffer risk more than ever
- regulations that limit the ability to diversify can have other considerable
  benefits but it's important to be aware of trade-offs inherent in these
  regulations when making management decisions
-->

<!--Our results provide additional support for the idea that individual fishers can
reduce their income variability by diversifying their catch portfolios. However,
most individuals do not change their catch diversity substantially from year to
year [OLE: is there a supplemental figure or statistic that speaks to this
point?], and most strategies adopted by Alaskan fishermen lie in a high-risk,
high-reward quadrant in which reducing diversification increases both revenue
variability and expected revenue. Previous work has suggested that changes to
fisheries management such as limited entry or catch share programs may
inadvertently act as barrier to diversification and promote specializing
(Hilborn et al. 2003; Kasperski and Holland 2013). While some of these
specialized fishing strategies may be associated with increased revenue
variability, the benefit may be that they are also associated with increased
revenue.-->

TODO: finish this:
External market forces may interact in concert with management regulations to
promote specialization for fishers or other individuals dependent on natural
resources. With increasing environmental variability and extreme events in
a changing climate (REF), it is particularly important that individuals who rely
on natural resources as a primary source of income have strategies to reduce
risk. Though variability in individual revenue may not be actively incorporated
into many management decisions, analyses like ours help to understand the range
of diversification strategies available and how their benefits may change over
time.

\bibliographystyle{pnas2011}
\bibliography{refs}

\clearpage

# Figures

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.7\textwidth]{../figs/fig1-jk.pdf}

\caption{Trends in revenue and diversification over time. (A) Map of fishing
regions in Alaska, USA. (B) Total revenue in billions of 2009 USD from `r
range_yrs[1]` to `r range_yrs[2]` for multiple (grey) and single (blue) permit
holders. (C) Proportion of single permit holders for all fisheries (thick black
line) and with permit holders omitted each year who fished any of the `r
n_jacknifed` top species groups (thin lines). (D) Mean effective diversity of
species fished (Simpson's diversity index) for all fisheries (thick black line)
and with permit holders omitted each year who fished any of the `r n_jacknifed`
top species groups (thin lines). } \label{fig:ts} \end{center} \end{figure}

\clearpage

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.7\textwidth]{../figs/stan-across-strategy-variability2.pdf}

\caption{Individuals who fish permits and permit combinations (permit
strategies) with higher species diversity have lower expected revenue
variability. On the y-axis, dots, thick lines, and thin lines represent
posterior medians, 50\%, and 95\% credible intervals for the `r nstr` most
common permit strategies. Thick black line, and dark and light grey shaded
regions indicate median, 50\%, and 95\% credible intervals of strategy-level
regression built into the hierarchical model. Estimated variability represents
expected standard deviation for a permit holder who does not change species
diversity or days fished from year-to-year. }

\label{fig:across}
\end{center}
\end{figure}


\clearpage

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.9\textwidth]{../figs/stan-gg-spoke2.pdf}

\caption{Diversifying permits reduces risk and usually increases revenue. Two
examples of expected revenue variability and median revenue for owners of (A)
individual halibut and (B) gillnet herring roe permits (dark blue) compared with
owners of the same permit plus one more (medium blue) or two more (light
blue) permits. On the y-axis, dots and lines represent posterior median and 95\%
credible intervals of expected standard deviations of year-to-year revenue rates
of return.}

\label{fig:spokes}
\end{center}
\end{figure}

\clearpage

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.6\textwidth]{../figs/stan-offset-break-anti-spaghetti-specializing.pdf}

\caption{Effect of specializing on year-to-year changes in revenue and revenue
variability for individual permit holders \emph{within} the `r nstr` most common
strategies. Changing species diversity from year-to-year is associated with
higher variability for individuals within the strategies above the zero line.
Specializing is associated with greater expected revenue ratios to the right of
the zero line and lower expected revenue ratios to the left of the zero line.
For example, individuals with setnet permits for salmon in Prince William Sound,
Cook Inlet, Kodiak, Bristol Bay, or the Alaska Peninsula (second purple dot from
right) target primarily sockeye salmon and experience a $\sim$1.5\% increase in
their year-to-year revenue ratio (horizontal axis) and a $\sim$0.5\% increase in
their revenue variability (vertical axis) for a 1\% decrease in their ratio of
year-to-year species diversity. These effects are estimated holding fishing
season length constant from year-to-year. Dots and line segments represent
medians and 50\% credible intervals of posteriors. The effect of generalizing
(increasing species diversity) is similar (Fig. S\ref{fig:mirror}).}

\label{fig:within}
\end{center}
\end{figure}

\clearpage

# Materials and methods

<!-- TODO: get fig:re in results -->
<!-- Our model estimates overall average effects (Fig. S\ref{fig:fe}) and effects -->
<!-- that vary by permit set (Fig. S\ref{fig:re}). -->

We obtained commercial fisheries landings and revenue data for all permit
holders in Alaska from `r range_yrs[1]` to `r range_yrs[2]` from the Alaskan
Commercial Fisheries Entry Commission (CFEC). These data represent a total of
nearly `r round(ndat$initial_tot_rev/1e9)` billion 2009 USD. We adjusted for
inflation by setting all revenues to 2009 USD [@bea2016]. We implemented
a number of filtering steps to focus on individuals actively engaged in
commercial fishing (Supporting Information). In particular, we removed permit
holders whose median revenue was less than \$10,000 2009 USD to focus on
individuals for whom commercial fishing formed a substantial part of their
income.

We used combinations of state and federal fishing permits held by an individual
in a specific year to define fishing "strategies", representing the species
targeted (Supporting Information). For example, we combined permits that were
otherwise the same except for vessel size (e.g. a longliner fishing sablefish
on a vessel under 60 ft was considered to have the same strategy as a longliner
fishing sablefish on a vessel over 60 ft), and we combined a number of permits
that were targeting a single species and differed only in region fished.
Therefore, for example, someone fishing herring roe in Kodiak and someone
fishing herring roe in Alaska Peninsula were both considered to have a "herring
roe" permit group. For salmon permits, we formed salmon strategies based on
similarities in gear type and species composition (Supporting Information).

After this process of combining permits, we were left with `r n_grouped_permits`
unique fishing permit groups (Table S\ref{tab:strategies}), which we aggregated
within each person-year combination to form permit strategies. For example, if
someone fished halibut and sablefish in the same year, their strategy for that
year would be "halibut-sablefish". Specific details on our strategy definitions
are available in the Supporting Information.

Our models focused on species diversity as a possible predictor of revenue and
revenue variability while controlling for effort. For consistency with previous
analyses [@vedeld2007; @illukpitiya2008; @debela2012; @kasperski2013;
sethi2014], we calculated effective species diversity for permit holder $i$ and
time (year) $t$ as the inverse of Simpson's diversity index $\lambda$
[@simpson1949; @jost2006] (also referred to in economics as the
Herfindahl-Hirschman Index) weighted by revenue $R$:

\begin{equation}
1/\lambda_{i,t} = 1 / \left(\sum\limits_{s=1}^{n_{i,t}}
  \left(R_{i,s,t}/ \sum\limits_{s=1}^{n_{i,t}} R_{i,s,t}\right)\right)^2,
\end{equation}

\noindent where $s$ indexes species from $1$ through $n$. We calculated effort
or effective season length as the sum of the season lengths for each permit
each year. Season length for a given permit was calculated as the number of
days between the first instance of fishing a permit and the last instance of
fishing a permit each year. For example, if someone fished for 30 days on a
salmon drift gillnet permit in Southeast Alaska and 40 days on a statewide
longline halibut permit, that individual's season length was calculated as 70
days.

## Hierarchical model

To jointly model revenue and revenue variability as a function of
individual-level covariates, we extended the basic variance-function regression
to a hierarchical Bayesian variance-function regression model with Gaussian
errors predicting $\log$ revenue, $\log(R_{i,t})$, for each fisher $i$ and time
$t$:

\begin{equation}
\log(R_{i,t}) \sim \mathrm{N} \left(\mu_{i,t}, \sigma_{i,t}^2 \right).
\end{equation}

\noindent Each year $t$ of revenue for fisher $i$ is assigned a strategy (permit
set) indexed by $j$ (Supporting Information, Table \ref{tab:strategies}) and
$\mu_{i,t}$ is modelled as:

\begin{equation}
\label{eq:mean}
\mu_{i,t} = \beta_{0,j,t} -
  \beta_{1,j} S_{i,t} I_{i,t} +
  \beta_{2,j} S_{i,t} I_{i,t} +
  \beta_3 D_{i,t} -
  \beta_4 S_{i,t} D_{i,t} I_{i,t} +
  \beta_5 S_{i,t} D_{i,t} I_{i,t} +
  \log(R_{i,t-1}),
\end{equation}

\noindent where $S_{i,t}$ is the log ratio of species diversity from
year-to-year, $\log \left(\mathrm{div}_{i,t}/\mathrm{div}_{i,t-1}\right)$,
$D_{i,t}$ is the log ratio of days fished from year-to-year, $\log
\left(\mathrm{days}_{i,t}/\mathrm{days}_{i,t-1}\right)$, and $I_{i,t}$ is an
indicator variable that takes a value $0$ if $\mathrm{div}_{i,t} \ge
\mathrm{div}_{i,t-1}$ (species diversity increasing; generalizing) and takes
a value $1$ if $\mathrm{div}_{i,t} < \mathrm{div}_{i,t-1}$ (species diversity
decreasing; specializing). Estimating a separate coefficient for the effect of
increasing diversity and decreasing diversity allows the costs and benefits of
these changes to be asymmetric.

We use $\log(R_{i,t-1}$) as an offset, and thus our model describes the log
ratio of revenue from year-to-year, i.e., $\log(R_{i,t}) - \log(R_{i,t-1})
= \log(R_{i,t}/R_{i,t-1})$. We model change in revenue because preliminary
analysis revealed non-stationary trends in revenue (resulting from experience or
other unmodeled factors). This approach eliminates having to deal with the
non-stationary time series and eliminates having to estimate thousands of
mean-revenue intercepts associated with individual fishers. We removed
person-year combinations without a subsequent year of revenue.

The $\beta$ terms represent estimated coefficients, some of which are allowed to
vary as "random effect" terms. The intercepts $\beta_{0,j,t}$ are allowed to
vary by strategy-year combinations, are centred on zero, and are constrained by
a normal distribution:

\begin{equation}
\beta_{0,j,t} \sim \mathrm{N} \left(0, \sigma^2_{\beta_{j,t}} \right).
\end{equation}

\noindent This approach, combined with the lack of a global intercept,
constrains the mean change in log revenue across all fishers and years to be
zero. I.e., we assume that the data are stationary after accounting for all
covariates and including an intercept that varies with strategy-year
combinations. Initial versions of the model confirmed that the global intercept
along with strategy-level intercepts would be estimated at almost exactly zero.
The estimated $\beta_{j,t}$'s represent years when people with a particular
permit set tended to collectively do better or worse than the long-term average
(Fig. S\ref{fig:str-yrs}). Therefore, an individual's expected change in revenue
(and variability) is after accounting for the general trend in revenue for that
permit set.

The slopes $\beta_{1,j}$ and $\beta_{1,j}$ are then allowed to vary by strategy
with estimated means:

\begin{align}
\beta_{1,j} &\sim \mathrm{N} \left(\mu_{\beta_1}, \tau^2_{\beta_1}\right),\\
\beta_{2,j} &\sim \mathrm{N} \left(\mu_{\beta_2}, \tau^2_{\beta_2}\right).
\end{align}

\noindent The $\beta_{1,j}$'s and $\beta_{2,j}$'s represent how much a 1\%
decrease and 1\% increase, respectively, in the ratio of species diversity from
year-to-year within a strategy translates to a given percent change in the ratio
of revenue from year-to-year. These are after controlling for changes in effort,
$\beta_3$, and the interaction between changes in effort and changes in species
diversity, $\beta_4$ and $\beta_5$.

Whereas traditional hierarchical linear models assume homoscedasticity --- that
the residual variance does not vary systematically --- our variance-function
regression model allows the residual variance to be heteroscedastic and vary as
a function of predictors. Instead of estimating a single residual error scale
term, $\sigma$, we model the scale of the residual error with another
hierarchical model with a similar form:

\begin{equation}
\label{eq:cv}
\sigma_{i,t} =
\exp \left( \gamma_{0,j} -
\gamma_{1,j} S_{i,t} I_{i,t} +
\gamma_{2,j} S_{i,t} I_{i,t} +
\gamma_3 D_{i,t} -
\gamma_4 S_{i,t} D_{i,t} I_{i,t} +
\gamma_5 S_{i,t} D_{i,t} I_{i,t}\right).
\end{equation}

\noindent Again, $S_{i,t}$ and $I_{i,t}$ represent, respectively, the log ratio
of species diversity from year-to-year and an indicator variable for increasing
or decreasing species diversity. We exponentiate the equation to ensure all
scale parameters, $\sigma_{i,t}$, are positive. Residuals from the mean
component of our model clearly show a "V"-shaped pattern, which the variance
component of our model can capture (Fig. S\ref{fig:resid}). Furthermore, the
downside residuals (year-individual data points with lower than expected
revenue) look qualitatively similar to the complete set of residuals (upside
and downside) (Fig. S\ref{fig:resid-downside}). Therefore, it seemed
appropriate to model the variability as a proxy for revenue risk.

The $\gamma$'s are estimated coefficients, three of which are allowed to vary by
strategy, constrained by normal distributions:

\begin{align}
\gamma_{0,j}  &\sim \mathrm{N} \left(\eta_0 + \eta_1 M_j + \eta_2 E_j, \tau^2_{\gamma_0}\right),\\
\gamma_{1,j}  &\sim \mathrm{N} \left(\mu_{\gamma_1}, \tau^2_{\gamma_1}\right),\\
\gamma_{2,j}  &\sim \mathrm{N} \left(\mu_{\gamma_2}, \tau^2_{\gamma_2}\right).
\end{align}

\noindent The $\gamma_{0,j}$ coefficients are modelled with two strategy-level
predictors. The symbols $M_j$ and $E_j$ represent the mean species diversity
and mean combined fishing season length (effort) for strategy $j$. The
coefficients $\eta_1$ and $\eta_2$ represent predictors in this strategy-level
regression estimating effects across strategies.

<!-- TODO: rerun with 2.14.1 -->

We fit our models with Stan 2.14.1 [@standevelopmentteam2016;
@standevelopmentteam2016a] and R 3.3.2 [@rcoreteam2016]. Stan implements the
No-U-Turn Hamiltonian Markov chain Monte Carlo algorithm [@hoffman2014] to
perform Bayesian statistical inference. We assigned weakly informative priors
on all parameters: $N(0, 2^2)$ priors on all $\beta$, $\gamma$, and $\eta$
parameters, and half-$t(3, 0, 2)$ priors (i.e. degrees of freedom of 3 and
scale of 2) on all $\sigma$ and $\tau$. We ran four chains and 2,000 iterations
and discarded the first 1,000 iterations of each chain as warm up. We checked
for chain convergence visually with trace plots, ensured that $\hat{R}<1.05$
(the potential scale reduction factor) and that the effective sample size, as
calculated by the rstan R package [@standevelopmentteam2016], was greater than
200 for all parameters [@gelman2014]. We verified that our model returned
sensible estimates by comparing our estimates to an ad-hoc two-stage model fit
with the lme4 R package [@bates2015], where we first fit a mixed-effects
regression model to the mean, and then modelled the absolute residuals from
that first model in a second mixed-effects model representing the variance
component. We also validated the model with simulated data, and were able to
recover the true values (Supporting Information). Code used in our analysis is
available at <https://github.com/NCEAS/pfx-commercial/portfolios/>.

# Acknowledgments

TODO. National Center for Ecological Analysis and Synthesis (NCEAS) working
group *Applying portfolio effects to the Gulf of Alaska ecosystem: Did
multi-scale diversity buffer against the Exxon Valdez oil spill?*. Other
participants... Smith...

<!-- vim: textwidth=80 -->
