---
title: 'Reduced diversity and increased specialization of Alaska salmon fishermen:
  effects on variation in individual revenues'
author: "Eric, Sean, Ole, Rich, Milo, Jordan, Jen, Alan, Ben, etc."
date: '2017-02-08'
output:
  html_document: default
  pdf_document:
    fig_caption: yes
  word_document: default
bibliography: CatchPortfolios_salmon.bib
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy=FALSE, tidy.opts=list(width.cutoff=60), warning = FALSE, message = FALSE)
library(knitr)
library(tidyverse)
library(reshape2)
library(date)
#library(ggtern)
library(mgcv)
#library(fpc)
library(gridExtra)
#library(glmmTMB)
library(zoo)
library(rstan)
library(ggrepel)
modelname = "model-linear" # model-linear-nointeractions
```

## Abstract
\setlength\parindent{24pt}

On a large scale, individual fishers may experience positive outcomes from diversifying their catches; examples include reduced year-to-year variability in annual revenue and increases in average revenue. Fishers can increase their species diversity by owning a permit that allows multiple species to be targeted, or by owning multiple permits. External forces such as management regulations (limited entry,  catch shares), the number of permits available, or market forces (permit or fish prices) may disincentive individuals to hold multiple permits. Thus, individuals holding a single permit may only be able to increase species diversity within the confines of their permit (longer fishing seasons, broader spatial area). Using a long-term dataset of individual salmon fishers in Alaska, we develop new models for understanding how the effects of diversification have changed through time. These changes allow us to compare patterns across regions and gear, as well as better understand how external forces have changed incentives for diversifying (hatchery production, market forces, EVOS?). 

## Introduction

[broader paragraph on diversification with extensions to terrestrial agriculture etc, ]

Previous analyses have largely focused on patterns of diversification across ecosystems and the benefits to adopt more or less diverse fishing strategies [@kasperski_income_2013; @anderson_benefits]. Fisheries in Alaska are one of the most productive seafood regions in the world (both by landed weight, and revenue [@office_of_science_and_technology_fisheries_2016], FAO cite) and they allow individuals to adopt a wide range of fishing strategies. [@anderson_benefits] defined strategies as a permit or collection of permits held by single individuals. While adding new permits allows fishers to increase species diversity, few individual permits allow individuals to harvest more than one species (examples of single-species permits in the Gulf of Alaska include halibut, sablefish, herring, crab, shellfish, shrimp, sea cucumber). Some exceptions to these single species permits include participants in federally managed groundfish fisheries and participants in state managed salmon fisheries.

Salmon fisheries are particularly important in Alaska both in terms of participation and value. The number of fishers participating only in salmon fisheries has increased over the last 30 years as a fraction of the total [@anderson_benefits] (figure XX) and overall participation, total value?, and median revenue has increased since the early 2000s (Fig. A1 and A2).  Yet the landed species diversity has declined over time (averaged across individuals, Fig. A6) and across different salmon fisheries there appears to be little correlation between diversity and revenue (Fig. A4). Thus the relationship between harvest diversity and revenue is neither simple or straight-forward. As seen in terrestrial agriculture[@purdy_financial_1997], diversification may affect the mean or CV of revenue, but not both, or the effect of diversification for revenue may vary over time [bad last sentence].

Salmon permits that allow the catch of multiple species provide an unusual opportunity for examining how the availablity of resources and fisher choice interact to affect both the total revenue and variability in revenue achieved by individual fishers. All individuals owning a single permit have the same opportunites to fish and to choose different fishing strategies. Individual fishers can therefore elect to utilize a diverse set of available species or be highly specialized on a single species. Thus the variability in fisher behavior among indivdiual fishers exist a range of strategies that allow us to understand the relationship between diversity of harvest and economic outcomes at the scale of the individual making decisions. Most previous analyses have focused on either vessel or community-level effects of diversification which are aggregate properties one step away from actual decision makers.  

Five species of salmon are harvested in Alaska waters (Chinook [*Oncorhynchus tshawytscha*], chum [*O. keta*], pink [*O. gorbuscha*], sockeye [*O. nerka*], coho [*O. kisutch*]), and the ability of individual salmon fishers to diversify across species is in part a function of geography and fishing gears. For example, salmon may be caught with purse seine gear (n = 6 areas), drift gillnets (5 areas), and set gillnets (12 areas). Other gears include beach seine and troll gears (though these are generally restricted to Kodiak and Southeast Alaska, respectively). Most of these gear types allow fishers the opportunity to harvest a diversity of salmon species.  In practice, though, the catch is dominated by a single species for some permits. Perhaps the best examples are the Bristol Bay set and drift gillnet fisheries where all five salmon species are present but fishers target sockeye salmon almost exclusively (REF). For these fisheries, specialization is associated with higher revenue and reduced CV in annual revenue [@anderson_benefits].

An additional, underappreciated consideration for understanding the relationship between diversification and fisheries value is how the availablity of different species in the environment affects the harvest diversity realized by individual fishers. For salmon one of the most imporant factors determining salmon abundance has been the expansion of artifical hatchery propagation of salmon species. Alaska’s large-scale hatchery program began in the late 1970s (Stopha 2015, Rich will get a reference), with substantial increases in hatchery production during the 1980s following low returns [@mahnken_historical_1998; @leber_stock_2008-1]. Many of these hatchery programs increased production to stabilize the variability of salmon returns [@hilborn_review_2001], though these programs may also have negative impacts on wild production [@wertheimer_relationship_2005; @ruggerone_productivity_2015]. The intra- and inter-specific effects of large scale hatchery programs has been debated [@hilborn_review_2001; @hilborn_hatcheries_1992; @wertheimer_comment:_2001; @heard_hatchery], but the effects of hatcheries have largely been focused on ecological interactions of aggregate yield to the fishery [@leber_stock_2008-1] or economic impacts [@boyce_alaska_1993], rather than to variability experienced by individual fishermen. 

[Jordan: Maybe a paragraph or notes about the importance of the role that salmon are serving in buffering or stabilizing fishing revenues in the state and how subsequently, understanding the responses of salmon to climate dynamics is especially critical (Mueter et al. 2002, and maybe Schindler et al., 2010)]

In addition to the changing species composition of fish available to be harvested, the benefits (or costs) of diversification are affected by a number of exogenous drivers, including the influx of farmed salmon, consolidation of salmon buyers, marketing, prices, new processing technology [@knapp_trends]. [Break out as own paragraph+add more examples?] Following price crashes in the 1990s and early 2000s (cite CFEC), many of these factors have been positive for salmon fishermen (increasing trends in prices and catches, Fig. A3, A12). In contrast, the late 1980s - early 1990s experienced high abundance and low prices (in part due to collapses of foreign currencies). Combined with the increased trend in hatchery production and resulting changes in species composition of adult salmon, these external factors may have also altered incentives for diversification. 

To understand the effects of diversification on revenue and revenue variability, we analyze single-permit salmon holders landings in Alaska between 1985 and 2015. We focus on XXX mixed species salmon fisheries in Alaska (ZZZ seine and YYY gillnet fisheries) and explore three aspects of the diversity-revenue relationship. First, we describe salmon catches - including total biomass, species composition, and diversity of catches -, fisheries participation, and revenue in Alaska over the past 30 years for each region and fishery. Second, we construct statistical time-series models to understand patterns of reveune and variability for each fishery. Specifically, we extend the methods used by [@anderson_benefits] to estimate time-varying effects of diversity on revenue and revenue variability. We then use estimated time-varying parameters to ask if the effects of diversity are constistent across time and similar among gear types and regions. Finally, we link our results to available information on economic and environmental data to uncover potential connections between the diversity-revenue relationship and longer term drivers in the region. 


*Ole's proposed flow of figures.(based on Eric's figure numbers circa 1//24/2017)
  + Generally, I would add Bristol Bay sockeye at least in the time-series figures because it is such a well known and well studied fishery.  It also would be a nice contrast to the others just to show that it has effectively no diversity at all.  It may not warrant inclusion in the modeling or later analysis, but I think not mentioning the most valuable salmon fishery in Alaska is a bit odd. Also, it should be mentioned that Bristol Bay is weird and not really representative of most salmon fisheries in Alaska.
  + A Suggest putting Appendix Figs 1,2,3 as a 3 panel main text figure- pick out the focal fisheries and plot trends in participation in one panel, trends in per-capita landed weight in a second, and trends in revenue in the third (may be best to put a smoother through all of the raw data to make it palatable to the eye).  Point is that despite being in different regions and different gears, there are broad similarities in all fisheries: falling participation, rising catches, relatively flat revenue.
  + B I like figure 1, but would add Bristol Bay (see above.). May need to switch to areas as rows, and the two panels as columns.  Also I would add a total catch line to the top panel (maybe add a third panel? not sure). 
  + C Fig 2 & 3.  Suggest making the percent composition plots one array of figures and having all 6 or 7 of the diversity plots in one figure with different colors or line-types. Maybe make the diversity figure big and then the composition for each area smaller and to the side? Main point is areas differ in composition and differ in overall diversity & trend in diversity. The diversity panel may be better served in the bullet point A, showing raw trends in diversity with other raw attributes.
  + D Fig. 4. Important points here seem to be about the variation across years (some gear-regions are more variable than others) and I wonder if this would be better presented as an esimated among-year variance or SD for each area and then possibly a pairwise correlations tile-plot among areas.  Suggest putting Fig. 4 time-series in the appendix.
  + E Fig 5 and 6.  Suggest collapsing multi-panel in Fig.5 to one panel and Fig.6 to one panel. Point of Fig. 5 is choosing to add species is negative on revenue and increasingly negative in recent years.  Story of Fig 6. is a little more confused, and no clear signal of diversity on revenue variability
  + Use final figure or two to relate external information to changes in diversity-revenue or diversity-revenue variability relationship.  Possibilities: total hatchery production, price of permits.  Also could break each area into revenue-variability bi-plots and plot each fishery in that bivariate space by time block (e.g. 5 year time blocks) 
  
Extra stuff:

Tradeoffs between specialization and diversification may also be found in case studies of terrestrial agriculture. For small scale farmers, increasing crop diversity may increase revenue and stabilize year-to-year income variability [@di_falco_crop_2003]. At larger scales or in the developed world, however, the incentives for diversification may be different. In the United States and Canada for example, crop diversity has declined since the late-1970s [@aguilar_crop_2015; @bradshaw_farm-level_2004]; some of the factors responsible include changing national policies, technology, and increased demand from other parts of the world (e.g. increased U.S. production of soybeans to meet demand from Asia). Shifts to from diverse of high value cash-crops to specializing on high-volume low-value crops such as cereals may have analogs for salmon fishermen in low value species with rapid life histories (pink and chum salmon).  


## Methods

### Data  

We obtained fisheries landings and revenue data for all iundividual permit holders in Alaska from 1985-2014 from the Commercial Fisheries Entry Commission (CFEC). Each permit holder has a unique identifier that allows us to track quantities including the species caught and revenue received. We adjusted for inflation by converting all revenues to 2009 USD. Because of our focus on salmon fishers, we were only interested in single permit fisheries, and we adopted a simplified version of the filtering steps described in [@anderson_benefits]. After aggregating individual trip-level catch and revenue to annual data, we retained permit holders who fished only one of the six salmon permits included in a year (Table 1, Fig. A5). Like [@anderson_benefits] we also worked with a log-differenced dataset to remove non-stationary individuals; to avoid complications of individuals switching strategies (e.g. changing from a seine fishery in Kodiak to a gillnet fishery in Bristol Bay), we only included pairs of years for a given person where they did not change strategies (i.e., didn’t change fisheries or add permits, etc.).  

```{r, echo=FALSE}
m = matrix("", 6, 3)
m[1,] = c("S01A", "Purse Seine", "Southeast Alaska")
m[2,] = c("S01E", "Purse Seine", "Prince William Sound")
m[3,] = c("S01K", "Purse Seine", "Kodiak")
m[4,] = c("S01M", "Purse Seine", "Alaska Peninsula")
m[5,] = c("S03A", "Gillnet", "Southeast Alaska")
m[6,] = c("S03E", "Gillnet", "Prince William Sound")
colnames(m) = c("Permit", "Gear", "Region")
kable(m)
```

### As hatchery programs have increased production, have individual salmon fishermen experienced reduced variability?  

Hatchery release trends for some regions included in our data have been summarized previously (e.g. [@ward_evaluating] summarized the trends for Prince William Sound). Release patterns have been dominated by pink and chum salmon ([@stopha2016]) and production has increased since the late 1970s (217 million juvenile pink salmon in PWS in 1985 versus 672 million in 2014). Though our analysis with CFEC data doesn't separate hatchery versus wild fish, as a contrast we can make comparisons across regions and species to examine potential effects of hatchery fish on variability. For example, we can compare the variation in the largely hatchery pink salmon returns (earnings and weight) from Prince William Sound to the largely wild pink salmon returns from Southeast Alaska. Similarly, we can compare the variability in earnings from species dominated by hatchery returns (pink, chum) to those with less hatchery influence (Chinook, sockeye). To make these comparisons, we aggregated the earnings and landings by species and permit, then calculated 3-year rolling CVs. As a second comparison, we examined whether long-term participants in any of the fisheries had experienced a change in variability. Similar to [@kasperski_income_2013] who calculated CVs of vessels before and after significant management events, we identified individuals who fished the same permit for at least three years in the first decade of our data (1985-1995) and at least three years in the last decade of the data (2004-2014). 

### Analysis 3: [Core of paper] Does effect of diversification change over time?

[@anderson_benefits] implemented a hierarchical Bayesian model that allowed group level predictors and covariates (diversity of catch, days fished) to simultaneously affect the mean and the variance of revenue for individual fishers. This model also included random intercepts by strategy and year; these terms allowed all permit holders in a given year to experience common fluctuations (external drivers including price changes, fluctuations in fish population abundance, or environmental disasters such as EVOS). 

To ask how benefits of diversification may have changed over time, we extend the model of [@anderson_benefits] to also include time-varying random effects in the slope coefficients of diversification. We define the variables $\Delta { S }_{ t }=\log\left( { S }_{ t }/{ S }_{ t-1 } \right)$ to represent the percent change in species diversity of the catch between years and $\Delta { D }_{ t }=\log\left( { D }_{ t }/{ D }_{ t-1 } \right)$ to represent the percent change in fishing effort (days fished) between years. The model for the expected log revenue of individual *i* in year *t*, $$\widehat { \log\left( { R }_{ i,t } \right)  } ={ B }_{ 0,j,t }+{ B }_{ 1,j,t }\cdot \Delta S_{ i,t }+{ B }_{ 2,j,t }\cdot \left( \Delta S_{ i,t }\cdot { S }_{ i,t-1 } \right) +{ B }_{ 3,j }\cdot \Delta D_{ i,t }+\log\left( { R }_{ i,t-1 } \right) $$, where *j* subscripts the salmon permit (Table 1). We model the time-varying permit specific intercepts as random and exchangeable across years, ${ B }_{ 0,j,t }\sim \mathrm{Normal}\left( { u }_{ { B }_{ 0 } },{ \sigma  }_{ { B }_{ 0 } } \right)$. The term ${ B }_{ 1,j,t }$ represents the permit level random effect of diversifying species. Our model extends that of [@anderson_benefits] to include the interaction terms $\Delta S_{ t }\cdot { S }_{ t-1 }$ which allow the effect of diversification to be dependent on the previous level of diversification. We assume this relationship is linear, and the coefficients ${ B }_{ 1,j,t }$ and ${ B }_{ 2,j,t }$ allow this to vary by permit and through time. The flexibility in these terms allows the benefit of increasing species diversity to be positive in some years (generalizing benefits revenue), negative in other years (specializing benefits revenue) and neutral in other years (revenue is independent of alterations to catch composition). We constrain the random effects associated with diversification to be modeled as autoregressive processes with shared variances across strategies, for example: ${ B }_{ 1,j,t }\sim \mathrm{Normal}\left( { B }_{ 1,j,t-1 },{ \sigma  }_{ { B }_{ 1 } } \right)$. Like [@anderson_benefits] we also allow changing effort (measured as percent change in days fished) ${ B }_{ 4,j }$ to vary by strategy, but we do not allow this effect to vary through time. Finally, the term $\log\left( { R }_{ i,t-1 } \right)$ is included as an offset [@anderson_benefits]. 
  Like [@anderson_benefits], we also included all of the above covariates in the variance model. We assumed the residuals from the mean model described above to be normally distributed, $$\log\left( { R }_{ i,t } \right) \sim \mathrm{Normal}\left( \widehat { \log\left( { R }_{ i,t } \right)  } ,\widehat { { \sigma  }_{ i,t } }  \right) $$. While conventional hierarchical linear regression assumes variances to be constant or shared by grouping variables, we also included fixed and random covariates, making the model akin to variance function regression or stochastic volatility models (CITE). Using the same notation as for the mean, we can express the log standard deviation as $$\widehat { \log\left( { \sigma  }_{ i,t } \right)  } ={ G }_{ 0,j }+{ G }_{ 1,j,t }\cdot \Delta S_{ i,t }+{ G }_{ 2,j,t }\cdot \left( \Delta S_{ i,t }\cdot { S }_{ i,t-1 } \right) +G_{ 3,j }\cdot \Delta D_{ i,t }$$. Unlike the linear model for the mean, we did not include an offset term. Further, we constrained the variance intercepts ${ G }_{ 0,j }$ to be temporally constant by strategy, instead of the dynamic nature of the intercepts in the mean model. Like the mean model, the random effects were constrained to be autoregressive, ${ G }_{ 1,j,t }\sim \mathrm{Normal}\left( { G }_{ 1,j,t-1 },{ \sigma  }_{ { G }_{ 1 } } \right)$, and covariates associated with days fished and species diversity were included as predictors. 

[Rich's note: This and delta D might be tricky for pink salmon stocks, which alternate between strong even or odd years (depends on the area) and may dictate fishery participation. For these, could we, should we, use log(St/St-2)??]

All done in Stan [@stan_development_team_stan_2015] using the NUTS algorithm [@hoffman_no-u-turn_2014].

Specific things that could be looked at: 
- EVOS (closure for Kodiak)  
- PWS pink salmon crash in 1992/1993: Pacific Salmon & their Ecosystems: Status and Future Options
- sockeye marketing 1983: [Jordan: So the Copper River salmon marketing ramped up in about 1983. I wonder if we could see a steady increase in value or movement towards sockeye in the early years of the data set] 

Does diversification exist within permits and if so, do we see similar advantages of diversification on revenue + CV?  
  
PWS chum and sockeye (increasing hatchery % over time) vs. Bristol Bay (all wild).  

CV of variability for high value species (sockeye, Chinook) vs. low value species (pink and chum). Coho fall somewhere in the middle. In other words, how have choices of species to produce by hatcheries affected the CV of revenue for individual permits? % hatchery fish of the harvest could be used as a continuous variable in the model.  
  
Similar to above, do life-history characteristics influence the variability of revenue? For example, sockeye and chum have substantially more complicated life-history traits (1-5 ocean years) than pink salmon (1 ocean year only). For high-volume hatcheries such as pink salmon: do annual returns affect price and revenue the following year more than for high-value fisheries?  
## Results

Our summaries of changing variability (aggregate for each fishery, across individuals, or within individuals) was not consistent with the hypothesis that variability in revenue or landings has decreased over time (Fig. A9-A10). For pink and chum salmon, fisheries in PWS generally experience higher year-to-year variability compared to SEAK. In looking at the changing CV of individuals, we see a slight increase in the CV of revenue for people fishing in the recent decade who were also fishing 1985-1995 (with the exception being S01A). 

Figure 1 and 2 shows the time series of species composition and effective spp diversity. Interesting patterns are (1) overall low diversity of S01E (pinks), (2) increasing chum in S03A

Figure 3 shows that one component of variability is the year-to-year variation that individuals experience as a result of external forces (stock abundance, prices, etc) as well as common fluctuations in revenue experienced by all permit holders (Anderson et al., n.d.). This shows the time varying intercepts on the mean -- negative equals losing revenue on average, while positive implies increasing revenue. The panels of this figure show big differences between permit types (gillnets tend to have lower year-to-year variation). Some permits have extreme variation; for example, S01E (PWS seine) has fluctuations between exp(-1) and exp(+1), corresponding to year-to-year changes of 36-270%.
  
Figure 4 shows the time varying nature of the model for all permits. These are the time series of benefits of changing from effective species diversity of 1 to 2, and here we are isolating year-to-year differences in revenue (the intercepts) from diversification benefits. In some cases the effect is positive (fishing more spp yields more revenue). But in many cases this is not the case, and the trend is negative (indicating that specialization is more profitable over time). In some cases, these time series mirror the aggregate diversity in the raw time series (Fig. 1). I think we also expect this result - fishermen increase diversity or not, because it's profitable.
  
Figure 5 shows that temporal trends, by permit, in going from 1 to 2 species fished on the variance (again removing the overall intercepts). From [@anderson_benefits] we'd expect diversification to reduce variance. A few examples of how to read this: S01A is slightly positive for most of the time series, but rapidly increases in the last 10 years -- this means generalists with diverse portfolios have had increases in year-to-year variability. For S01E, the start and end of the time series are associated with these 'costs' to diversification. For S03E, we see a more gradual trend (cost of diversification increases).

## Discussion

- In terms of participation, why did people drop out in 2000 in SEAK and Alaska Peninsula, but not in PWS? PWS largely seems like a boutique fishery on the Copper River -- which includes a qualtiy component (Copper River at forefront) and a timing issue relative to the rest of the Gulf of Alaska populations (buffers against market impacts). In addition, there are many indications that costs have increased over time as a result of requirements by fish buyers for fish of higher quality (refrigerated sea water, ice, increased handling time, more expensive vessels) more expensive permits, etc.  

- Revenue is going up since 2000 -- but each permit holder is fishing more (more days, more total pounds) and the trend in weight is faster than the trend in price.  

- Ability to specialize primarily driven by many forces beyond control of fishermen. But in some situations, diversification is possible by extending the fishing season to exploit additional spp (fishing earlier for chum in PWS S01E). Or fishing different areas to exploit more spp in S03 (SEAK).  

- Prices are generally synchronous for all spp (especially coho/chum/pink), and have increased at a faster rate compared to chum and pink salmons since 2000 (Fig A8-A9). We're also seeing a trend toward greater synchrony between species (Fig A11-A12, A14). Projecting future prices is notably complex [@knapp_trends]. Synchrony in prices may largely be function of marketing, improved product, freezing, etc.  

- We're seeing increased specialization within permits. People in PWS are trending toward increased specialization on pink salmon. People in SEAK trending toward increased specialization on chum. Both of these are switches from higher value species (sockeye, chum) to lower value species. Both pink and chum have higher variability in year to year returns than sockeye.  

- One potential exogenous driver related to management that can be ruled out as being responsible for increased specialization of salmon fishermen is changes in the timing or length of fishing seasons. Shifting seasons to later in the year may reduce catch of early run fish for example. However, we found no significant trend in either the season length or average season date for any fishery (Fig. A15). Also, management actions over time have generally tried to eliminate interception fisheries. Examples include the closure of the South Kayak Island area where sockeye bound for Cook Inlet were being harvested off the Copper River. Rich can provide some citations. 

- The decisions to implement a hatchery program for a given species has economic consequences for overall revenue and the variability of that revenue to fishers and other participants. Salmon that have complex life histories in the ocean phase of their life cycle (sockeye and possibly chum salmon) may have an inherent portfolio that buffers against large annual swings in abundance compared to pink and coho salmon. Also, high-value species may experience less annual fluctuation than lower value species.  

- For small scale farmers, economists have often wondered why farmers don't diversify more. Diversification can be profitable, but in the short term risky or expensive. Bigger than cost seems to be the risk [@karlan_agricultural_2012]

- The potential to diversify may be limited for areas with a limited number of available species or by managements actions that prohibit fishing in certain areas (e.g., PWS where seiners are generally confined to areas/times with pink and chum salmon whereas gillnet fishers are able to fish in areas with sockeye, chum, Chinook, coho, and pink salmon). Rich can provide a citation for the PWS management plan.  

- One of the outputs that can be calculated from the raw data is variability of revenue over time (e.g. year to year, at the level of individual & permit). This gets at more of the social impacts of changing opportunities, but long term trends in rising income variability have been documented in the US (Hacker 2006). Understanding trends in revenue of individuals whose incomes are dependent on natural resources are interesting because they tend to be more variable than that of the larger population [@mishra_income_2002]. Are there specific attributes that correlate with rising variability (for example the CBO has suggested variability in year to year income is negatively correlated with income [@cbo_recent_2008]) 

### Robustness of strategies to changing environments [stash in appendix?]

One of the challenges in modeling the effects of species diversification for individual fishers is that between two years, both the species composition of the catch and the species composition of the fish available to be caught change (making it hard to disentangle the selectivity of fishermen versus portfolio of fish available). To isolate the effects of a changing environment from changing strategies, we used a subset of the data from fishery permits where discrete strategies were emergent. For details see Appendix. 

Using purse seine permit holders in Prince William Sound as an example, we first calculated the species composition (weighted by revenue) for each active permit holder in each year. For each pair of years that an individual fished, we calculated the squared Euclidian distance between the species compositions of those years, $$\Delta { p }_{ ind,t }=\sum _{ sp=1 }^{ 5 }{ { \left| { p }_{ ind,sp,t }-{ p }_{ ind,sp,t-1 } \right|  }^{ 2 } } $$.  

Larger values of $\Delta { p }_{ ind,t }$ indicate greater year-to-year discrepancy in species composition. To focus on individuals who had more constant strategies in adjacent years, we ommitted data corresponding to the upper third of this distribution (retaining $\Delta { p }_{ ind,t }$ < 0.334). Next, we used the species composition from each individual and pair of years to cluster individuals into strategies. Clustering was performed using the partitioning around medoids (pam) algorithm in R with an estimated number of clusters (using 'pamk' in fpc library), and selected 2 groups of individual strategies: pink salmon specialization (medoid proportion of pink salmon = 0.95), and a strategy focused on pink salmon and chum (medoid proportions of 0.64 and 0.31 respectively). For each group, we then calculated the mean and CV of revenue; we calculated the log difference of revenue for each individual, $\Delta { R }_{ i }=\log\left( { R }_{ i,t+1 }/{ R }_{ i,t } \right)$ and then calculated the CV as the standard deviation over people, $sd(\Delta { R }_{ i=1:n })$.  

\break  

```{r, echo=FALSE, fig.pos="placeHere", fig.cap="CV and mean annual revenue by salmon permit, 1975-2014. Average participation (unique fishers) given by circle size, and effective species diversity is shown with color shading.", fig.height=3.2, eval=FALSE}
dat.annual = readRDS(file="../../portfolio/data-generated/salmon-annual-for-modeling.rds")

cfec_all = readRDS("../../portfolio/data-generated/salmon_data_for_paper.rds")
people = group_by(cfec_all, p_fshy, year) %>% 
  summarize(n = length(unique(p_holder))) %>% 
  group_by(p_fshy) %>% summarize(n = mean(n,na.rm=T))
names(people)[which(names(people)=="p_fshy")] = "strategy_permit"
people$strategy_permit = paste0(substr(people$strategy_permit,1,1), substr(people$strategy_permit,3,5))

dat.annual = filter(dat.annual, npermit==1 & substr(strategy_permit,1,1)=="S") %>%
  group_by(strategy_permit) %>% 
  mutate(nPeople = length(unique(p_holder))) %>% 
  filter(nPeople > 200) %>% 
  dplyr::select(-nPeople) %>% 
  group_by(strategy_permit, year) %>% 
  summarize(meanrev = mean(revenue), meandiv = mean(specDiv)) %>% 
  arrange(strategy_permit, year) %>% 
  group_by(strategy_permit) %>% 
  summarize(mean_rev = mean(meanrev,na.rm=T), cv_rev = sd(diff(log(meanrev)),na.rm=T),
    diversity = mean(meandiv,na.rm=T)) %>% as.data.frame
  
dat.annual = left_join(dat.annual, people)
dat.annual$gear = substr(dat.annual$strategy_permit,2,3)
dat.annual$gear[which(dat.annual$gear=="01")]="Purse seine"
dat.annual$gear[which(dat.annual$gear=="03")]="Drift gillnet"
dat.annual$gear[which(dat.annual$gear=="04")]="Set gillnet"
dat.annual$gear[which(dat.annual$gear=="05")]="Hand troll"
dat.annual$gear[which(dat.annual$gear=="15")]="Power troll"

dat.annual$price = NA
dat.annual$price[dat.annual$strategy_permit=="S01A"]=250000
dat.annual$price[dat.annual$strategy_permit=="S01E"]=186700
dat.annual$price[dat.annual$strategy_permit=="S01H"]=84800
dat.annual$price[dat.annual$strategy_permit=="S01K"]=40000
dat.annual$price[dat.annual$strategy_permit=="S01L"]=227500
dat.annual$price[dat.annual$strategy_permit=="S01M"]=56900

dat.annual$price[dat.annual$strategy_permit=="S03A"]=88900
dat.annual$price[dat.annual$strategy_permit=="S03E"]=224200
dat.annual$price[dat.annual$strategy_permit=="S03H"]=63500
dat.annual$price[dat.annual$strategy_permit=="S03M"]=119500
dat.annual$price[dat.annual$strategy_permit=="S03T"]=148200
dat.annual$price[dat.annual$strategy_permit=="S03W"]=NA

dat.annual$price[dat.annual$strategy_permit=="S04D"]=88900
dat.annual$price[dat.annual$strategy_permit=="S04H"]=15300
dat.annual$price[dat.annual$strategy_permit=="S04K"]=77000
dat.annual$price[dat.annual$strategy_permit=="S04M"]=55900
dat.annual$price[dat.annual$strategy_permit=="S04T"]=38500
dat.annual$price[dat.annual$strategy_permit=="S04W"]=7300
dat.annual$price[dat.annual$strategy_permit=="S04X"]=5400
dat.annual$price[dat.annual$strategy_permit=="S04Y"]=9900
dat.annual$price[dat.annual$strategy_permit=="S04Z"]=11100

dat.annual$price[dat.annual$strategy_permit=="S05B"]=11000
dat.annual$price[dat.annual$strategy_permit=="S15B"]=38300

g1 = ggplot(dat.annual, aes(x=log(mean_rev), y=log(cv_rev), label = dat.annual$strategy_permit)) + geom_point(aes(col=gear,size=diversity,alpha=0.7)) + scale_size(range = c(0, 20)) + geom_text_repel() + ylab("CV annual revenue") + xlab("Mean revenue (dollars)") + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
print(g1)
pdf("Figure_01.pdf")
print(g1)
dev.off()
```


```{r, echo=FALSE, fig.pos="placeHere", fig.cap="Harvest of hatchery produced salmon through time and by area.", fig.height=3.2}

seak_hatch = read.csv("../../salmon/data-generated/seak_hatcheryHarvest.csv")
pws_hatch = read.csv("../../salmon/data-generated/pws_hatcheryHarvest.csv")
kodiak_hatch = read.csv("../../salmon/data-generated/kodiak_hatcheryHarvest.csv")

seak_hatch = melt(seak_hatch, id.vars = c("Year"))
pws_hatch = melt(pws_hatch, id.vars = c("Year"))
kodiak_hatch = melt(kodiak_hatch, id.vars = c("Year"))

seak_hatch$region = "Southeast"
pws_hatch$region = "Prince William Sound"
kodiak_hatch$region = "Kodiak"

seak_hatch$value[is.na(seak_hatch$value)] = 0
pws_hatch$value[is.na(pws_hatch$value)] = 0
kodiak_hatch$value[is.na(kodiak_hatch$value)] = 0

dat = bind_rows(seak_hatch, pws_hatch)
dat = bind_rows(dat, kodiak_hatch)

names(dat)[which(names(dat)=="variable")]="species"
dat$value = dat$value / 1000000

# chinook chum coho pink sock
cbPalette <- c("#009E73", "#4B2E83", "#999999", "#CC79A7", "#D55E00")

dat$col = as.numeric(as.factor(dat$species))
g1 = ggplot(dat, aes(Year, value)) +
  geom_area(aes(fill = species), position = "stack") +
  xlab("Year") + ylab("Harvest of hatchery produced salmon (millions)") +
  scale_fill_manual(values = c('Chinook'=cbPalette[1], 'Chum'=cbPalette[2], 
    'Coho'=cbPalette[3], 'Pink'=cbPalette[4], 'Sockeye'=cbPalette[5]), name="Species",
    labels =c("Chinook","Sockeye","Coho","Pink","Chum")) + 
  facet_wrap(~region, scale="free_y")

# bring in total harvest data (hatchery + wild)
hatch = read.csv("../../salmon/data-generated/commonPropertyCommercialCatch.csv")
hatch = hatch[,c("Year","Chinook","Sockeye","Coho","Pink","Chum","Region")]
hatch$Region = as.character(hatch$Region)
hatch$Region[which(hatch$Region=="PWS")] = "Prince William Sound"
hatch = melt(hatch, id.vars = c("Year","Region"))
names(hatch)[which(names(hatch)=="variable")]="Species"

g2 = ggplot(hatch, aes(Year, value, color=Species)) + 
  geom_line() + facet_wrap(~Region) + 
  xlab("Year") + ylab("Hatchery contribution to catch (%)") + 
  scale_color_manual(values = c('Chinook'=cbPalette[1], 'Chum'=cbPalette[2], 
    'Coho'=cbPalette[3], 'Pink'=cbPalette[4], 'Sockeye'=cbPalette[5]), name="Species",
    labels =c("Chinook","Sockeye","Coho","Pink","Chum"))

gridExtra::grid.arrange(g1, g2)

```



```{r, echo=FALSE, fig.pos="placeHere", fig.cap="Time varying species composition and diversity, by permit. Species are colored as follows: green (Chinook), purple (chum), silver  (coho), pink (pink), red (sockeye).", fig.height=3.2}
#cfec <- feather::read_feather("../../portfolio/data-generated/salmon.feather")
#saveRDS(cfec, "portfolio/data-generated/salmon_data_for_paper.rds")
cfec = readRDS("../../portfolio/data-generated/salmon_data_for_paper.rds")
cfec$year = as.numeric(substr(cfec$landdate, 1, 4))

# plot proportion of species landed by salmon permit

g1 = group_by(cfec[cfec$p_fshy=="S 01A" & cfec$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, p_fshy) %>%
  summarize(p = sum(g_earn)) %>% group_by(year, p_fshy) %>% mutate(p = p/sum(p), col = as.numeric(as.factor(spec))) %>%
  ggplot(aes(year, p, group=spec, color = spec)) + geom_line() +  
  ylab("Species %") + xlab("") + ggtitle("S01A") + 
scale_colour_manual(values = c('CHNK'=cbPalette[1], 'CHUM'=cbPalette[2], 
    'COHO'=cbPalette[3], 'PINK'=cbPalette[4], 'SOCK'=cbPalette[5]), name="Species",
    labels =c("Chinook","Chum","Coho","Pink","Sockeye")) + theme(legend.position="none")

g2 = group_by(cfec[cfec$p_fshy=="S 01E" & cfec$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, p_fshy) %>%
  summarize(p = sum(g_earn)) %>% group_by(year, p_fshy) %>% mutate(p = p/sum(p), col = as.numeric(as.factor(spec))) %>%
  ggplot(aes(year, p, group=spec, color = spec)) + geom_line() +  
  ylab("Species %") + xlab("") + ggtitle("S01E") + theme(legend.position="none") + 
scale_colour_manual(values = c('CHNK'=cbPalette[1], 'CHUM'=cbPalette[2], 
    'COHO'=cbPalette[3], 'PINK'=cbPalette[4], 'SOCK'=cbPalette[5]))

g3 = group_by(cfec[cfec$p_fshy=="S 01K" & cfec$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, p_fshy) %>%
  summarize(p = sum(g_earn)) %>% group_by(year, p_fshy) %>% mutate(p = p/sum(p), col = as.numeric(as.factor(spec))) %>%
  ggplot(aes(year, p, group=spec, color = spec)) + geom_line() +  
  ylab("Species %") + xlab("") + ggtitle("S01K") + theme(legend.position="none") + 
scale_colour_manual(values = c('CHNK'=cbPalette[1], 'CHUM'=cbPalette[2], 
    'COHO'=cbPalette[3], 'PINK'=cbPalette[4], 'SOCK'=cbPalette[5]))

g4 = group_by(cfec[cfec$p_fshy=="S 01M" & cfec$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, p_fshy) %>%
  summarize(p = sum(g_earn)) %>% group_by(year, p_fshy) %>% mutate(p = p/sum(p), col = as.numeric(as.factor(spec))) %>%
  ggplot(aes(year, p, group=spec, color = spec)) + geom_line() +  
  ylab("Species %") + xlab("") + ggtitle("S01M") + theme(legend.position="none") + 
scale_colour_manual(values = c('CHNK'=cbPalette[1], 'CHUM'=cbPalette[2], 
    'COHO'=cbPalette[3], 'PINK'=cbPalette[4], 'SOCK'=cbPalette[5]), name = "Species", labels =c("Chinook","Chum","Coho","Pink","Sockeye")) + theme(legend.position="none")

g5 = group_by(cfec[cfec$p_fshy=="S 03A" & cfec$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, p_fshy) %>%
  summarize(p = sum(g_earn)) %>% group_by(year, p_fshy) %>% mutate(p = p/sum(p), col = as.numeric(as.factor(spec))) %>%
  ggplot(aes(year, p, group=spec, color = spec)) + geom_line() +  
  ylab("Species %") + xlab("") + ggtitle("S03A") + theme(legend.position="none") + 
scale_colour_manual(values = c('CHNK'=cbPalette[1], 'CHUM'=cbPalette[2], 
    'COHO'=cbPalette[3], 'PINK'=cbPalette[4], 'SOCK'=cbPalette[5]))

g6 = group_by(cfec[cfec$p_fshy=="S 03E" & cfec$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK") & cfec$year > 0,], year, spec, p_fshy) %>%
  summarize(p = sum(g_earn)) %>% group_by(year, p_fshy) %>% mutate(p = p/sum(p), col = as.numeric(as.factor(spec))) %>%
  ggplot(aes(year, p, group=spec, color = spec)) + geom_line() +  
  ylab("Species %") + xlab("") + ggtitle("S03E") + theme(legend.position="none") + 
scale_colour_manual(values = c('CHNK'=cbPalette[1], 'CHUM'=cbPalette[2], 
    'COHO'=cbPalette[3], 'PINK'=cbPalette[4], 'SOCK'=cbPalette[5]))

dat.diff = readRDS(file="../../portfolio/data-generated/cfec-diff-for-modeling.rds")
dat.diff = dat.diff[which(dat.diff$revenue >= 10000 & dat.diff$revenue.prev >= 10000), ]
dat.salmon = dat.diff[dat.diff$strategy_permit%in%c("S01A", "S01E", "S01K", "S03A", "S03E","S01M"),]

h1 = group_by(dat.salmon[dat.salmon$strategy_permit=="S01A",], year, strategy_permit) %>%
  summarize(meanDiv = mean(specDiv), low = quantile(specDiv,0.25), upp = quantile(specDiv,0.75)) %>%
  ggplot(aes(year, meanDiv)) + geom_line() + 
  geom_ribbon(aes(ymin= (low), ymax= (upp)), alpha=0.4) + ylab("Diversity") + xlab("") + theme(legend.position="none")

h2 = group_by(dat.salmon[dat.salmon$strategy_permit=="S01E",], year, strategy_permit) %>%
  summarize(meanDiv = mean(specDiv), low = quantile(specDiv,0.25), upp = quantile(specDiv,0.75)) %>%
  ggplot(aes(year, meanDiv)) + geom_line() + 
  geom_ribbon(aes(ymin= (low), ymax= (upp)), alpha=0.4) + ylab("Diversity") + xlab("") + theme(legend.position="none")

h3 = group_by(dat.salmon[dat.salmon$strategy_permit=="S01K",], year, strategy_permit) %>%
  summarize(meanDiv = mean(specDiv), low = quantile(specDiv,0.25), upp = quantile(specDiv,0.75)) %>%
  ggplot(aes(year, meanDiv)) + geom_line() + 
  geom_ribbon(aes(ymin= (low), ymax= (upp)), alpha=0.4) + ylab("Diversity") + xlab("") + theme(legend.position="none")

h4 = group_by(dat.salmon[dat.salmon$strategy_permit=="S01M",], year, strategy_permit) %>%
  summarize(meanDiv = mean(specDiv), low = quantile(specDiv,0.25), upp = quantile(specDiv,0.75)) %>%
  ggplot(aes(year, meanDiv)) + geom_line() + 
  geom_ribbon(aes(ymin= (low), ymax= (upp)), alpha=0.4) + ylab("Diversity") + xlab("") + theme(legend.position="none")

h5 = group_by(dat.salmon[dat.salmon$strategy_permit=="S03A",], year, strategy_permit) %>%
  summarize(meanDiv = mean(specDiv), low = quantile(specDiv,0.25), upp = quantile(specDiv,0.75)) %>%
  ggplot(aes(year, meanDiv)) + geom_line() + 
  geom_ribbon(aes(ymin= (low), ymax= (upp)), alpha=0.4) + ylab("Diversity") + xlab("") + theme(legend.position="none")

h6 = group_by(dat.salmon[dat.salmon$strategy_permit=="S03E",], year, strategy_permit) %>%
  summarize(meanDiv = mean(specDiv), low = quantile(specDiv,0.25), upp = quantile(specDiv,0.75)) %>%
  ggplot(aes(year, meanDiv)) + geom_line() + 
  geom_ribbon(aes(ymin= (low), ymax= (upp)), alpha=0.4) + ylab("Diversity") + xlab("") + theme(legend.position="none")

gridExtra::grid.arrange(g1, g2, g3, h1, h2, h3, nrow=2)
```

```{r, echo=FALSE, fig.pos="placeHere", fig.cap="Time varying species composition and diversity, by permit.", fig.height=3.2} 
gridExtra::grid.arrange(g4, g5, g6, h4, h5, h6, nrow=2)
```
  
\break  

```{r, eval=FALSE,echo=FALSE}
load("../../salmon/analysis/cv-model-simple.Rdata")
# groups here are 80% pink vs 97% pink

yrs = 1986:2014

hist(extract(mod)[["g_offset"]], 40, col="grey", xlab="Effect of specializing on pink (sd)")

xx = exp(extract(mod)[["sd_yr"]])
df.gen = data.frame("year"=yrs, "mean" = apply(xx, 2, mean), "low" = apply(xx, 2, quantile, 0.025),
  "hi" = apply(xx, 2, quantile, 0.975), "strategy"="pink:chum")

offset = matrix(extract(mod)[["g_offset"]],ncol=1) %*% matrix(rep(1,length(yrs)), nrow=1)
xx = exp(extract(mod)[["sd_yr"]] + offset)
df.spec = data.frame("year"=yrs, "mean" = apply(xx, 2, mean), "low" = apply(xx, 2, quantile, 0.025),
  "hi" = apply(xx, 2, quantile, 0.975), "strategy"="pink")

ggplot(data=rbind(df.gen,df.spec), aes(year, mean, group=strategy, fill=strategy)) + geom_ribbon(aes(ymin=low,ymax=hi), alpha=0.5) + geom_hline(yintercept=0, color="red")

```


```{r, echo=FALSE, fig.pos="placeHere", fig.cap="Time varying intercepts on estimated mean revenue, by permit. These are interpreted as the mean change in revenue for each permit holder in each year (positive equates to increasing revenue, etc), after accounting for effects of catch diversification and days fished."}
# grab original data to look at permit names
diffdat = readRDS(file="../../salmon/data-generated/salmon.rds")

years = seq(min(diffdat$year), max(diffdat$year))
permits = levels(as.factor(diffdat$permit))

# note that we have to use factor() here instead of unique() because
# that's the same as the ordering in the estimation code
yearpermit = data.frame("yrpermit"=levels(as.factor(diffdat$year_permit)))
yearpermit$year = as.numeric(substr(yearpermit$yrpermit,1,4))
yearpermit$permit = substr(yearpermit$yrpermit,6,9)

yearpermitold=yearpermit
load("../../salmon/analysis/model-linear.Rdata")
yearpermit=yearpermitold
est = extract(mod, permute="TRUE")

yearpermit$b0 = apply(est$b0_str_yr,2,median)
yearpermit$b0_low = apply(est$b0_str_yr,2,quantile,0.025)
yearpermit$b0_hi = apply(est$b0_str_yr,2,quantile,0.975)

ggplot(data=yearpermit) + 
  geom_ribbon(aes(x=year,ymin=b0_low,ymax=b0_hi),alpha=0.3) +
  geom_line(aes(x=year,y=b0)) + 
  facet_wrap(~permit) + xlab("Year") + 
  ylab("Intercepts on mean") + 
  geom_hline(yintercept=0, color="red")
```
  
\break  
  
```{r, echo=FALSE}
# fig.pos="placeHere", fig.cap="Time varying effects of adding 1 species on revenue (measured in ln dollars) for purse seine permit in Prince William Sound (S01E). This figure illustrates two things: (1) the benefit of adding more species diminishes as a function of past diversity, and (2) the economic benefit to adding more species has been reduced through time."
# grab original data to look at permit names
diffdat = readRDS(file="../../salmon/data-generated/salmon.rds")
years = seq(min(diffdat$year), max(diffdat$year))
#permits = levels(as.factor(diffdat$permit))

load("../../salmon/analysis/model-linear.Rdata")
yearpermit=yearpermitold

est = extract(mod, permute="TRUE")

# start with making time series of estimates on mean
# challenge is that they're time varying quadratic functions
# do facet_wrap of estimates by year

# effect of adding more 1 species from 1. specdiv.prev = 1

specdiv.prev = seq(1, 5, by=0.1)
specdiv = specdiv.prev + 1 # always add 1 more in these comparisons
x1 = log(specdiv/specdiv.prev)
x2 = log(specdiv/specdiv.prev) * specdiv.prev

df = expand.grid("year"=1:29, "diversity" = seq(1,length(x1)), "permit"=1:6)
df$permits = permits[df$permit]
df$years = years[df$year]
df$x = specdiv.prev[df$diversity]
df$pred = 0
df$low = 0
df$high = 0
b0_indx = match(paste(as.character(df$years), as.character(df$permits)), as.character(yearpermit$yrpermit)) 

for(i in 1:nrow(df)) {
  tmp = x1[df$diversity[i]] * (est$b_1[,df$year[i], df$permit[i]])# + x2[df$diversity[i]] * (est$b_2[,df$year[i], df$permit[i]]+est$b2_str_yr_mu)
  df$pred[i] = median(tmp)
  df$low[i] = quantile(tmp,0.025)
  df$high[i] = quantile(tmp,0.975)
}
df_s01e_b1 = df
#ggplot(df[df$permits=="S01E",], aes(x=x,y=pred,group=years)) +
#  #geom_ribbon(aes(ymin=low,ymax=high), alpha=0.4) +
#  geom_line() +
#  facet_wrap(~years) + xlab("Previous species diversity") + ylab("Predicted #benefit of adding +1 spp (ln $)")

```
  
\break  

```{r, fig.pos="placeHere", fig.cap="Time varying effects of adding 1 species on revenue (measured in ln dollars) for hypothetical individuals with species diversity ~ 1. This shows declining benefits of diversification for most fisheries, S01M.", echo=FALSE}
# grab original data to look at permit names
diffdat = readRDS(file="../../portfolio/data-generated/salmon-diff-for-modeling.rds")
years = seq(min(diffdat$year), max(diffdat$year))
#permits = levels(as.factor(diffdat$permit))

yearpermitold=yearpermit
load("../../salmon/analysis/model-linear.Rdata")
yearpermit=yearpermitold

est = extract(mod, permute="TRUE")

# start with making time series of estimates on mean
# challenge is that they're time varying quadratic functions
# do facet_wrap of estimates by year

# effect of adding more 1 species from 1. specdiv.prev = 1

specdiv.prev = seq(1, 5, by=0.1)
specdiv = specdiv.prev + 1 # always add 1 more in these comparisons
x1 = log(specdiv/specdiv.prev)
x2 = log(specdiv/specdiv.prev) * specdiv.prev
x3 = log(specdiv/specdiv.prev) * specdiv.prev^2

df = expand.grid("year"=1:length(years), "diversity" = seq(1,length(x1)), "permit"=1:6)
df$permits = permits[df$permit]
df$years = years[df$year]
df$x = specdiv.prev[df$diversity]
df$pred = 0
df$low = 0
df$high = 0
b0_indx = match(paste(as.character(df$years), as.character(df$permits)), as.character(yearpermit$yrpermit)) 

for(i in 1:nrow(df)) {
  #tmp = est$b0_str_yr[,b0_indx[i]] + 
  tmp = 
    x1[df$diversity[i]] * (est$b_1[,df$year[i], df$permit[i]])# +
    #x2[df$diversity[i]] * (est$b_2[,df$year[i], df$permit[i]]+est$b2_str_yr_mu)# +
    #x3[df$diversity[i]] * (est$b_3[,df$year[i], df$permit[i]]+est$b3_str_yr_mu)
  df$pred[i] = mean(tmp)
  df$low[i] = quantile(tmp,0.025)
  df$high[i] = quantile(tmp,0.975)
}

ggplot(df[df$diversity==1,], aes(x=years,y=pred,group=permits)) +
  geom_line() +
  geom_ribbon(aes(ymin=low,ymax=high), alpha=0.4) +
  facet_wrap(~permits) + xlab("Year") + ylab("Predicted benefit of adding +1 spp (ln $)") + geom_hline(yintercept=0, color="red")

```
  
\break  
  
```{r, fig.pos="placeHere", fig.cap="Time varying effects of adding 1 species on log standard deviation of revenue for hypothetical individuals with species diversity ~ 1.", echo=FALSE}
# grab original data to look at permit names
diffdat = readRDS(file="../../portfolio/data-generated/salmon-diff-for-modeling.rds")
years = seq(min(diffdat$year), max(diffdat$year))
#permits = levels(as.factor(diffdat$permit))

load("../../salmon/analysis/model-linear.Rdata")
yearpermit=yearpermitold

est = extract(mod, permute="TRUE")

# start with making time series of estimates on mean
# challenge is that they're time varying quadratic functions
# do facet_wrap of estimates by year

# effect of adding more 1 species from 1. specdiv.prev = 1

specdiv.prev = seq(1, 5, by=0.1)
specdiv = specdiv.prev + 1 # always add 1 more in these comparisons
x1 = log(specdiv/specdiv.prev)
x2 = log(specdiv/specdiv.prev) * specdiv.prev
x3 = log(specdiv/specdiv.prev) * specdiv.prev^2

df = expand.grid("year"=1:length(years), "diversity" = seq(1,length(x1)), "permit"=1:6)
df$permits = permits[df$permit]
df$years = years[df$year]
df$x = specdiv.prev[df$diversity]
df$pred = 0
df$low = 0
df$high = 0

for(i in 1:nrow(df)) {
  #tmp = est$g0 + est$g0_strategy[,df$permit[i]] + 
  tmp = 
    x1[df$diversity[i]] * (est$g_1[,df$year[i], df$permit[i]])# +
    #x2[df$diversity[i]] * (est$g_2[,df$year[i], df$permit[i]]+est$g2_str_yr_mu)
  df$pred[i] = median(tmp)
  df$low[i] = quantile(tmp,0.025)
  df$high[i] = quantile(tmp,0.975)
}

ggplot(df[df$diversity==1,], aes(x=years,y=pred,group=permits)) +
  geom_line() +
  geom_ribbon(aes(ymin=low,ymax=high), alpha=0.4) +
  facet_wrap(~permits) + xlab("Year") + ylab("Predicted marginal benefit of adding +1 spp (ln $)") + geom_hline(yintercept=0, color="red")

``` 
  
```{r, echo=FALSE, fig.pos="placeHere", fig.cap = "As model validation, we can compare the estimated intercepts for PWS S01E (purse seine) representing average year to year change in revenue versus the change year over year in ln(pink salmon returns). Correlation between them is ~ 0.8. This suggests (as expected) people on average make a lot more money in years when pink salmon returns are high. [side note: diversification may only be beneficial here when returns and prices are both low.]"}

yearpermit$b0 = apply(est$b0_str_yr,2,median)
yearpermit$b0_low = apply(est$b0_str_yr,2,quantile,0.025)
yearpermit$b0_hi = apply(est$b0_str_yr,2,quantile,0.975)

S01E = yearpermit[yearpermit$permit=="S01E",]
names(S01E)[which(names(S01E)=="year")]="Year"

returns = readxl::read_excel("../../salmon/paper/salmon_returns_PWS.xlsx")
pinkPWS = returns[,c("Year", "Pink_total", "Chum_total")]
pinkPWS = pinkPWS[which(pinkPWS$Year >= 1975),]
pinkPWS$diff = c(NA, diff(log(pinkPWS$Pink_total)))

pinkPWS = dplyr::left_join(pinkPWS, S01E)

ggplot(data=pinkPWS, aes(b0, diff,color=Year)) + geom_point() + stat_smooth(method=lm) + xlab("Estimated intercept from model") + ylab("Change in total pink salmon returns")

# bring in slopes from above
#slopes = df_s01e_b1[df_s01e_b1$permits=="S01E" & df_s01e_b1$diversity==1,]
#slopes = slopes[,c("years","pred")]
#names(slopes)[which(names(slopes)=="years")] = "Year"
#names(slopes)[which(names(slopes)=="pred")] = "B1"
#pinkPWS = dplyr::left_join(pinkPWS, slopes)

#pinkPWS$diffchum = c(NA, diff(log(pinkPWS$Chum_total)))

#prices = group_by(cfec[which(cfec$p_fshy%in%c("S 01E","S 03E") & #cfec$spec%in%c("PINK","CHUM","SOCK")),], year, spec) %>% 
#  summarize(meanPrice = mean(g_price)) %>%
#  arrange(year)
#prices <- dcast(melt(prices, id.vars = c("spec", "year")), year ~ #variable+spec)
#names(prices)[which(names(prices)=="year")]="Year"

#pinkPWS = dplyr::left_join(pinkPWS, prices)
#pinkPWS$pricediffchum = c(NA, diff(pinkPWS$meanPrice_CHUM))
#pinkPWS$pricediffpink = c(NA, diff(pinkPWS$meanPrice_PINK))
#pinkPWS$pricediffsock = c(NA, diff(pinkPWS$meanPrice_SOCK))

#cor(pinkPWS[,-c(5,6)], use="pairwise.complete.obs")

#ggplot(data=pinkPWS, aes(b0, B1)) + geom_point() + stat_smooth(method=lm) + xlab("Estimated intercept from model") + ylab("Change in total pink salmon returns")

#diversity = dplyr::group_by(dat.salmon[dat.salmon$strategy_permit=="S01E",], year) %>%
#  summarize(meanDiv = mean(specDiv))
#names(diversity)[which(names(diversity)=="year")] = "Year"
#pinkPWS = dplyr::left_join(pinkPWS, diversity)

```
  
  
\break  
  
  
# References
